---
title: "Testing the effect of source climate and genetic PCs on adaptive traits"
author: Anoob Prakash
lightbox: true
format: 
  html:
    code-overflow: scroll
    toc: true
    embed-resources: true
    code-tools: false
    code-fold: false
knitr:
  opts_knit:
    root.dir: "Z:/Anoob/Polygenic_Adaptation/ANALYSIS"
editor_options: 
  chunk_output_type: console
---
```{r}
#| message: false
#| warning: false
#| echo: false

# set working directory  
setwd("Z:/Anoob/Polygenic_Adaptation/ANALYSIS")
```

# Background

The following analysis is carried out to sparse out the effect of climate and admixture ancestry on locally adaptive traits in red spruce. The analysis is carried out for all the traits that was observed to be selected for under the Q~PC~ analysis.

**Steps**

-   Scatterplot: Visualizes the relationship between variables to check for non-linearity.

-   Model fitting: Compares a linear model with and without transformation.

-   AIC comparison: Quantifies which model fits better.

-   Diagnostic plots: Checks for normality and homoscedasticity of residuals.

-   QQ plot: Specifically checks residual normality.

```{r}
#| message: false
#| warning: false

# Packages loaded 
require(tidyverse)
require(kableExtra)
require(formattable)
require(lmerTest)
require(tidyverse)
require(sjPlot)
require(FactoMineR)
require(factoextra)
require(tidymodels)
require(modelsummary)
require(DT)
require(GGally)

```

## Data

```{r}
set.seed(123)

# Meta data
Meta <- read.table("../../MCMCglmm/Exome/RS_Exome_metadata.txt", sep="\t",header=T)

# Trait and plasticity
# data <- read.csv("../TraitData/trait_blups_complete.csv", header=T)

data <- read.csv("data/Qpc_data.csv", header=T)

# peek into the data
data |> 
  DT::datatable(
    caption = "Qpc Trait data",
    filter="top") |> 
  DT::formatRound(2:5, digits = 2)
```

### Climate data

```{r}
# Climate data
climateNA <- read.csv("../ClimateData/climateData_evolution.csv")
climateNA <- subset(climateNA,select=-c(X)) # remove overalapping columns
# climateNA <- climateNA[,-c(1)] # remove the Family column
# climateNA <- group_by(climateNA, Population) %>% summarise_if(is.numeric,mean)


# merge towaRDS THE END
# data <- merge(climateNA,traitBlups[,1:22],by="Family") # merge data by the column "Population"
# data$Region <- as.factor(data$Region); data$Region <- factor(data$Region,levels = c("Core", "Margin", "Edge"))

```

**Garden climate**

```{r garden climate and climate PCA}
Garden_clim <- read.table("../ClimateData/paper1_climate_data/climNA_gardens_2019-20.txt", header=T)

Garden_clim |> 
  DT::datatable(
    caption = "Garden climate",
    filter="top") |> 
  DT::formatRound(1:4, digits = 1)


colnames(Garden_clim)[1] <- "Garden"
colnames(Garden_clim)[2] <- "Year"

### estimate the mean climate variables for 2019-2020 across three garden sites
# Subset the data for MD
md_data <- Garden_clim %>% filter(Garden == "MD")
# Calculate column means for numeric columns
md_numeric <- colnames(md_data[-c(1:2)])
md_means <- apply(md_data[, md_numeric], 2, mean)
# Add the means as a new row
md_data_with_means <- rbind(md_data, c(Garden = "MD", Year = "Mean", md_means))

# Subset the data for NC
nc_data <- Garden_clim %>% filter(Garden == "NC")
# Calculate column means for numeric columns
nc_numeric <- colnames(nc_data[-c(1:2)])
nc_means <- apply(nc_data[, nc_numeric], 2, mean)
# Add the means as a new row
nc_data_with_means <- rbind(nc_data, c(Garden = "NC", Year = "Mean", nc_means))

# Subset the data for VT
vt_data <- Garden_clim %>% filter(Garden == "VT")
# Calculate column means for numeric columns
vt_numeric <- colnames(vt_data[-c(1:2)])
vt_means <- apply(vt_data[, vt_numeric], 2, mean)
# Add the means as a new row
vt_data_with_means <- rbind(vt_data, c(Garden = "VT", Year = "Mean", vt_means))

# combined mean data - added to the garden climate data.frame
Garden_clim <- rbind(Garden_clim,md_data_with_means[3,],nc_data_with_means[3,],vt_data_with_means[3,])

# Reset row names and revert columns to numeric
rownames(Garden_clim) <- NULL
Garden_clim[vt_numeric] <- lapply(Garden_clim[vt_numeric], as.numeric)
```

## Merge trait and climate data

```{r}
# combine the trait and climate data
data_clim <- left_join(data,climateNA[,-c(2:8)])
```

# Functions for Model testing

::: panel-tabset
## Function to run the models

```{r}
scale_models <- function(data, 
                          # variable to run the model
                          response, 
                          variable_predictors, # climate variables
                          constant_predictor,  # ancestry score - q2
                          # scale the response variable
                          scale_response = FALSE,
                          # scale the predictors
                          scale_predictors = FALSE) {
                          df <- data
                          all_predictors <- unique(c(variable_predictors, constant_predictor))
                          
                          # Optionally scale response
                          if (scale_response) {
                            df[[response]] <- scale(df[[response]])
                            response_var <- response
                          } else {
                            response_var <- response
                          }
                          
                          # Optionally scale predictors
                          if (scale_predictors) {
                            for (pred in all_predictors) {
                              df[[pred]] <- scale(df[[pred]])
                            }
                            var_preds <- variable_predictors
                            const_pred <- constant_predictor
                          } else {
                            var_preds <- variable_predictors
                            const_pred <- constant_predictor
                          }
                          
                          models_list <- list()
                          for (vp in var_preds) {
                            # Build formula string with interaction
                            formula_str <- paste0(response_var, " ~ ", 
                                                  vp, " + ", const_pred, " + ", vp, ":", const_pred)
                            model_formula <- as.formula(formula_str)
                            models_list[[vp]] <- lm(model_formula, data = df)
                          }
                          return(models_list)
                        }
```

## Function to summarize models

```{r}
summarize_models <- function(model_list, 
                                    coef_names = NULL,
                                    gof_stats = c("r.squared", "adj.r.squared"),
                                    star_levels = c(' ' = 0.05, '**' = 0.0045, '***' = 0.001),
                                    output_format = "default") {
  
                    # Ensure Intercept is included
                    if (!is.null(coef_names) && !"(Intercept)" %in% names(coef_names)) {
                      coef_names <- c("(Intercept)" = "Intercept", coef_names)
                    }
                    
                    modelsummary::modelsummary(
                      model_list,
                      estimate = "{estimate} ({std.error})",
                      statistic = "{p.value}{stars}",
                      coef_map = coef_names,
                      gof_map = gof_stats,
                      stars = star_levels,
                      output = output_format
                    )}

# name the coef for the models
## for Genetic PC1
coef_map_PC1 <- c(
  "CMD" = "Climate Variable",
  "DD18" = "Climate Variable",
  "EXT" = "Climate Variable",
  "MAR" = "Climate Variable",
  "MSP" = "Climate Variable",
  "PAS" = "Climate Variable",
  "PET" = "Climate Variable",
  "RH" = "Climate Variable",
  "DD_0" = "Climate Variable",
  "TD" = "Climate Variable",
  "eFFP" = "Climate Variable",
  "PC1" = "Genetic PC",
  "CMD:PC1" = "Climate:Genetic PC",
  "DD18:PC1" = "Climate:Genetic PC",
  "EXT:PC1" = "Climate:Genetic PC",
  "MAR:PC1" = "Climate:Genetic PC",
  "MSP:PC1" = "Climate:Genetic PC",
  "PAS:PC1" = "Climate:Genetic PC",
  "PET:PC1" = "Climate:Genetic PC",
  "RH:PC1" = "Climate:Genetic PC",
  "DD_0:PC1" = "Climate:Genetic PC",
  "TD:PC1" = "Climate:Genetic PC",
  "eFFP:PC1" = "Climate:Genetic PC"
)
## for Genetic PC2
coef_map_PC2 <- c(
  "CMD" = "Climate Variable",
  "DD18" = "Climate Variable",
  "EXT" = "Climate Variable",
  "MAR" = "Climate Variable",
  "MSP" = "Climate Variable",
  "PAS" = "Climate Variable",
  "PET" = "Climate Variable",
  "RH" = "Climate Variable",
  "DD_0" = "Climate Variable",
  "TD" = "Climate Variable",
  "eFFP" = "Climate Variable",
  "PC2" = "Genetic PC",
  "CMD:PC2" = "Climate:Genetic PC",
  "DD18:PC2" = "Climate:Genetic PC",
  "EXT:PC2" = "Climate:Genetic PC",
  "MAR:PC2" = "Climate:Genetic PC",
  "MSP:PC2" = "Climate:Genetic PC",
  "PAS:PC2" = "Climate:Genetic PC",
  "PET:PC2" = "Climate:Genetic PC",
  "RH:PC2" = "Climate:Genetic PC",
  "DD_0:PC2" = "Climate:Genetic PC",
  "TD:PC2" = "Climate:Genetic PC",
  "eFFP:PC2" = "Climate:Genetic PC"
)
```

## Function for 3D plot

```{r}
plot_in3d <- function(data, x, y, z, shape=NA, pch, grid.lines = 20, theta = 40, phi = 20,
                      xlab = "x-axis", ylab = "y-axis", zlab = "z-axis", color_var,
                      shape_legend_position=NA, show_legends=TRUE) {
  require(plot3D)
  require(paletteer)
  
  # Extract variables from the data frame
  var1 <- data[[x]] # Climate variable
  var2 <- data[[y]] # Ancestry
  var3 <- data[[z]] # trait 
  colvar <- data[[color_var]] # color 
  
  # Compute the linear regression
  fit <- lm(var3 ~ var1 + var2 + var1:var2)
  
  # Predict values on regular xy grid
  x.pred <- seq(min(var1, na.rm = TRUE), max(var1, na.rm = TRUE), length.out = grid.lines)
  y.pred <- seq(min(var2, na.rm = TRUE), max(var2, na.rm = TRUE), length.out = grid.lines)
  xy <- expand.grid(x = x.pred, y = y.pred)
  z.pred <- matrix(predict(fit, newdata = data.frame(var1 = xy$x, var2 = xy$y)), 
                   nrow = grid.lines, ncol = grid.lines)
  
  # Fitted points for droplines to surface
  fitpoints <- predict(fit)
  
  # plot shapes
  if(!is.na(shape)) {
    shape_vector <- as.numeric(factor(data[[shape]]))
    pch_values <- pch  # You can adjust these values for different shapes
    point_shapes <- pch_values[shape_vector]
  } else {
    point_shapes <- 16
  }
  
  # Scatter plot with regression plane
  scatter3D(var1, var2, var3, cex = 2, 
            theta = theta, # rotate the plot left/right direction
            phi = phi,     # rotate the plot top/bottom direction
            ticktype = "detailed",
            bty="b2", box=TRUE,
            colvar = colvar,
            col = paletteer_c("ggthemes::Sunset-Sunrise Diverging", 30, direction = 1),
            xlab = xlab, ylab = ylab, zlab = zlab, 
            pch = point_shapes,
            surf = list(x = x.pred, y = y.pred, z = z.pred,  
                        facets = NA, fit = fitpoints, col=c("grey70","black")),
            colkey = if(show_legends) list() else FALSE)
  
  # Update axis title font sizes
  # par(cex.lab=axis_title_font_size) 
  
  # Add shape legend if shape is specified and legends are to be shown
  if (!is.na(shape) && show_legends) {
    legend(shape_legend_position,
           legend = levels(factor(data[[shape]])),
           pch = pch[1:length(unique(data[[shape]]))],
           title = shape,
           cex = 0.8)
  }
}
```
:::

```{r}
# Climate model variables
climate_vars <- c('DD_0','DD18','MAR','PAS','MSP','RH','EXT','CMD','TD','eFFP','PET')
```

```{r}
#| warning: true
#| message: true

scatter_data <- data_clim
scatter_data[,climate_vars] <- lapply(scatter_data[climate_vars], scale)
scatter_data <- scatter_data %>% pivot_longer(cols = climate_vars, names_to = "ClimateVariables", values_to = "ClimateValue")
```

# Correlations between Climate, Ancestry and Kinship PCs

```{r}
#| fig-height: 15
#| fig-width: 15

set.seed(123)
pairs(data_clim[,c('DD_0','DD18','MAR','PAS','MSP','RH','EXT','CMD','TD','eFFP','PET','PC1','PC2','q2')])

# correlation df
corr_data <- data_clim[,c('DD_0','DD18','MAR','PAS','MSP','RH','EXT','CMD','TD','eFFP','PET','PC1','PC2','q2')]

names(corr_data)[14] <- "BS Ancestry"

corr_pair <- ggpairs(scale(corr_data))

corr_pair + theme_bw()

# ggsave("./figures/paper_figures/FigureS5_corr_plot.png", width = 3500, height = 3000,dpi=300,
#        limitsize = F,
#        units = "px", type = "cairo")
# 
# ggsave("./figures/paper_figures/FigureS5_corr_plot.pdf", width = 30, height = 25,dpi=300,
#        units = "cm")
```

# Height Growth

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(Growth))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
growth_models_unscaled <- scale_models(data = data_clim,
                                            response = "Growth",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
growth_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "Growth",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
growth_models_all_scaled <- scale_models(data = data_clim,
                                            response = "Growth",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(growth_models_unscaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(growth_models_pred_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(growth_models_all_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

growth_models_tinytables <- summarize_models(growth_models_all_scaled,
                                             coef_names = coef_map_PC2,
                                             output_format = "tinytable")
# tinytable::save_tt(growth_models_tinytables, "./tables/growth_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(growth_models_unscaled)

growth_aic_unscaled <- sapply(growth_models_unscaled, AIC)
growth_aic_pred_scaled   <- sapply(growth_models_pred_scaled, AIC)
growth_aic_all_scaled   <- sapply(growth_models_all_scaled, AIC)

growth_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = growth_aic_unscaled,
                              AIC_Pred_Scaled = growth_aic_pred_scaled,
                              AIC_All_Scaled = growth_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
growth_model_comparison <- growth_model_comparison[order(growth_model_comparison$AIC_Unscaled), ]

growth_model_comparison |> 
  DT::datatable(
    caption = "Growth Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)
```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(growth_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/growth_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(growth_models_unscaled[["PAS"]])

plot(growth_models_unscaled[["PET"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(growth_models_pred_scaled[["PAS"]])

plot(growth_models_pred_scaled[["PET"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(growth_models_all_scaled[["PAS"]])

plot(growth_models_all_scaled[["PET"]])
```
:::
:::::

------------------------------------------------------------------------

# Total Dry Biomass

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(TotalDry))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
tdm_models_unscaled <- scale_models(data = data_clim,
                                            response = "TotalDry",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
tdm_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "TotalDry",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
tdm_models_all_scaled <- scale_models(data = data_clim,
                                            response = "TotalDry",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(tdm_models_unscaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(tdm_models_pred_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(tdm_models_all_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

total_dry_models_tinytables <- summarize_models(tdm_models_all_scaled,
                                             coef_names = coef_map_PC2,
                                             output_format = "tinytable")

# tinytable::save_tt(total_dry_models_tinytables, "./tables/total_dry_biomass_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(tdm_models_unscaled)

tdm_aic_unscaled <- sapply(tdm_models_unscaled, AIC)
tdm_aic_pred_scaled   <- sapply(tdm_models_pred_scaled, AIC)
tdm_aic_all_scaled   <- sapply(tdm_models_all_scaled, AIC)

tdm_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = tdm_aic_unscaled,
                              AIC_Pred_Scaled = tdm_aic_pred_scaled,
                              AIC_All_Scaled = tdm_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
tdm_model_comparison <- tdm_model_comparison[order(tdm_model_comparison$AIC_Unscaled), ]

tdm_model_comparison |> 
  DT::datatable(
    caption = "Total Dry Biomass Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)

```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(tdm_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/total_dry_biomass_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(tdm_models_unscaled[["DD18"]])

plot(tdm_models_unscaled[["eFFP"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(tdm_models_pred_scaled[["DD18"]])

plot(tdm_models_pred_scaled[["eFFP"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(tdm_models_all_scaled[["DD18"]])

plot(tdm_models_all_scaled[["eFFP"]])
```
:::
:::::

------------------------------------------------------------------------

# Height Growth plasticity

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(Growth_plasticity))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
growth_plasticity_models_unscaled <- scale_models(data = data_clim,
                                            response = "Growth_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
growth_plasticity_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "Growth_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
growth_plasticity_models_all_scaled <- scale_models(data = data_clim,
                                            response = "Growth_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(growth_plasticity_models_unscaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(growth_plasticity_models_pred_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(growth_plasticity_models_all_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

growth_plasticity_models_tinytables <- summarize_models(growth_plasticity_models_all_scaled,
                                             coef_names = coef_map_PC2,
                                             output_format = "tinytable")
# tinytable::save_tt(growth_plasticity_models_tinytables, "./tables/growth_plasticity_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(growth_plasticity_models_unscaled)

growth_plasticity_aic_unscaled <- sapply(growth_plasticity_models_unscaled, AIC)
growth_plasticity_aic_pred_scaled   <- sapply(growth_plasticity_models_pred_scaled, AIC)
growth_plasticity_aic_all_scaled   <- sapply(growth_plasticity_models_all_scaled, AIC)

growth_plasticity_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = growth_plasticity_aic_unscaled,
                              AIC_Pred_Scaled = growth_plasticity_aic_pred_scaled,
                              AIC_All_Scaled = growth_plasticity_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
growth_plasticity_model_comparison <- growth_plasticity_model_comparison[order(growth_plasticity_model_comparison$AIC_Unscaled), ]

growth_plasticity_model_comparison |> 
  DT::datatable(
    caption = "Growth plasticity Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)

```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(growth_plasticity_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/growth_plasticity_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(growth_plasticity_models_unscaled[["PET"]])

plot(growth_plasticity_models_unscaled[["PAS"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(growth_plasticity_models_pred_scaled[["PET"]])

plot(growth_plasticity_models_pred_scaled[["PAS"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(growth_plasticity_models_all_scaled[["PET"]])

plot(growth_plasticity_models_all_scaled[["PAS"]])
```
:::
:::::

------------------------------------------------------------------------

# Total Dry Biomass plasticity

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(TotalDry_plasticity))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
tdm_plasticity_models_unscaled <- scale_models(data = data_clim,
                                            response = "TotalDry_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
tdm_plasticity_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "TotalDry_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
tdm_plasticity_models_all_scaled <- scale_models(data = data_clim,
                                            response = "TotalDry_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC2",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(tdm_plasticity_models_unscaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(tdm_plasticity_models_pred_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(tdm_plasticity_models_all_scaled,
                 coef_names = coef_map_PC2,
                 output_format = "gt")

tdm_plasticity_models_tinytables <- summarize_models(tdm_plasticity_models_all_scaled,
                                             coef_names = coef_map_PC2,
                                             output_format = "tinytable")

# tinytable::save_tt(tdm_plasticity_models_tinytables, "./tables/total_dry_biomass_plasticity_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(tdm_plasticity_models_unscaled)

tdm_plasticity_aic_unscaled <- sapply(tdm_plasticity_models_unscaled, AIC)
tdm_plasticity_aic_pred_scaled   <- sapply(tdm_plasticity_models_pred_scaled, AIC)
tdm_plasticity_aic_all_scaled   <- sapply(tdm_plasticity_models_all_scaled, AIC)

tdm_plasticity_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = tdm_plasticity_aic_unscaled,
                              AIC_Pred_Scaled = tdm_plasticity_aic_pred_scaled,
                              AIC_All_Scaled = tdm_plasticity_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
tdm_plasticity_model_comparison <- tdm_plasticity_model_comparison[order(tdm_plasticity_model_comparison$AIC_Unscaled), ]

tdm_plasticity_model_comparison |> 
  DT::datatable(
    caption = "Total Dry Biomass plasticity Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)

```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(tdm_plasticity_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/total_dry_biomass_plasticity_model_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(tdm_plasticity_models_unscaled[["DD18"]])

plot(tdm_plasticity_models_unscaled[["eFFP"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(tdm_plasticity_models_pred_scaled[["DD18"]])

plot(tdm_plasticity_models_pred_scaled[["eFFP"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(tdm_plasticity_models_all_scaled[["DD18"]])

plot(tdm_plasticity_models_all_scaled[["eFFP"]])
```
:::
:::::

------------------------------------------------------------------------

# Bud set (2019)

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(BudSet_2019))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
budset_19_models_unscaled <- scale_models(data = data_clim,
                                            response = "BudSet_2019",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
budset_19_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "BudSet_2019",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
budset_19_models_all_scaled <- scale_models(data = data_clim,
                                            response = "BudSet_2019",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(budset_19_models_unscaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(budset_19_models_pred_scaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(budset_19_models_all_scaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")

budset_19_models_tinytables <- summarize_models(budset_19_models_all_scaled,
                                             coef_names = coef_map_PC1,
                                             output_format = "tinytable")

# tinytable::save_tt(budset_19_models_tinytables, "./tables/budset_19_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(budset_19_models_unscaled)

budset_19_aic_unscaled <- sapply(budset_19_models_unscaled, AIC)
budset_19_aic_pred_scaled   <- sapply(budset_19_models_pred_scaled, AIC)
budset_19_aic_all_scaled   <- sapply(budset_19_models_all_scaled, AIC)

budset_19_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = budset_19_aic_unscaled,
                              AIC_Pred_Scaled = budset_19_aic_pred_scaled,
                              AIC_All_Scaled = budset_19_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
budset_19_model_comparison <- budset_19_model_comparison[order(budset_19_model_comparison$AIC_Unscaled), ]

budset_19_model_comparison |> 
  DT::datatable(
    caption = "Bud set (2019) Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)

```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(budset_19_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/budset_19_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(budset_19_models_unscaled[["DD_0"]])

plot(budset_19_models_unscaled[["PAS"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(budset_19_models_pred_scaled[["DD_0"]])

plot(budset_19_models_pred_scaled[["PAS"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(budset_19_models_all_scaled[["DD_0"]])

plot(budset_19_models_all_scaled[["PAS"]])
```
:::
:::::

------------------------------------------------------------------------

# Bud break (2020)

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(Budbreak_2020))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
budbreak_20_models_unscaled <- scale_models(data = data_clim,
                                            response = "Budbreak_2020",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
budbreak_20_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "Budbreak_2020",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
budbreak_20_models_all_scaled <- scale_models(data = data_clim,
                                            response = "Budbreak_2020",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(budbreak_20_models_unscaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(budbreak_20_models_pred_scaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(budbreak_20_models_all_scaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")

budbreak_20_models_tinytables <- summarize_models(budbreak_20_models_all_scaled,
                                             coef_names = coef_map_PC1,
                                             output_format = "tinytable")

# tinytable::save_tt(budbreak_20_models_tinytables, "./tables/budbreak_20_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(budbreak_20_models_unscaled)

budbreak_20_aic_unscaled <- sapply(budbreak_20_models_unscaled, AIC)
budbreak_20_aic_pred_scaled   <- sapply(budbreak_20_models_pred_scaled, AIC)
budbreak_20_aic_all_scaled   <- sapply(budbreak_20_models_all_scaled, AIC)

budbreak_20_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = budbreak_20_aic_unscaled,
                              AIC_Pred_Scaled = budbreak_20_aic_pred_scaled,
                              AIC_All_Scaled = budbreak_20_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
budbreak_20_model_comparison <- budbreak_20_model_comparison[order(budbreak_20_model_comparison$AIC_Unscaled), ]

budbreak_20_model_comparison |> 
  DT::datatable(
    caption = "Bud break (2020) Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)

```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(budbreak_20_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/budbreak_20_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(budbreak_20_models_unscaled[["DD18"]])

plot(budbreak_20_models_unscaled[["CMD"]])

plot(budbreak_20_models_unscaled[["eFFP"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(budbreak_20_models_pred_scaled[["DD18"]])

plot(budbreak_20_models_pred_scaled[["CMD"]])

plot(budbreak_20_models_pred_scaled[["eFFP"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(budbreak_20_models_all_scaled[["DD18"]])

plot(budbreak_20_models_all_scaled[["CMD"]])

plot(budbreak_20_models_all_scaled[["eFFP"]])
```
:::
:::::

------------------------------------------------------------------------

# Bud break plasticity (2020)

::::: panel-tabset
## Scatter plot

```{r}
# Scatter plot to check non linearity
ggplot(scatter_data, aes(x = ClimateValue, y = scale(BudBreak_cGDD_plasticity))) +
  geom_point() + facet_wrap(.~ClimateVariables, scales = "free_x") + theme_bw()
```

## Model summaries

```{r}
#| code-fold: true

# Variable Scaling
### Unscaled data
budbreak_20_plasticity_models_unscaled <- scale_models(data = data_clim,
                                            response = "BudBreak_cGDD_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = F,
                                            scale_predictors = F)


### Scaled data  
## Only the predictors are scaled
budbreak_20_plasticity_models_pred_scaled <- scale_models(data = data_clim,
                                            response = "BudBreak_cGDD_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = F,
                                            scale_predictors = T)

## Both the response and predictors are scaled
budbreak_20_plasticity_models_all_scaled <- scale_models(data = data_clim,
                                            response = "BudBreak_cGDD_plasticity",
                                            variable_predictors = climate_vars,
                                            constant_predictor = "PC1",
                                            scale_response = T,
                                            scale_predictors = T)
```

::: panel-tabset
### Unscaled data

```{r}
#| code-fold: true
summarize_models(budbreak_20_plasticity_models_unscaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")

```

### Predictors scaled

```{r}
#| code-fold: true
summarize_models(budbreak_20_plasticity_models_pred_scaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")
```

### All variables scaled

```{r}
#| code-fold: true
summarize_models(budbreak_20_plasticity_models_all_scaled,
                 coef_names = coef_map_PC1,
                 output_format = "gt")

budbreak_20_plasticity_models_tinytables <- summarize_models(budbreak_20_plasticity_models_all_scaled,
                                             coef_names = coef_map_PC1,
                                             output_format = "tinytable")

# tinytable::save_tt(budbreak_20_plasticity_models_tinytables, "./tables/budbreak_20_plasticity_model.docx")
```
:::

## Compare models using AIC

```{r}
#| code-fold: true
# Compare models using AIC
model_names <- names(budbreak_20_plasticity_models_unscaled)

budbreak_20_plasticity_aic_unscaled <- sapply(budbreak_20_plasticity_models_unscaled, AIC)
budbreak_20_plasticity_aic_pred_scaled   <- sapply(budbreak_20_plasticity_models_pred_scaled, AIC)
budbreak_20_plasticity_aic_all_scaled   <- sapply(budbreak_20_plasticity_models_all_scaled, AIC)

budbreak_20_plasticity_model_comparison <- data.frame(
                              # Model = model_names,
                              AIC_Unscaled = budbreak_20_plasticity_aic_unscaled,
                              AIC_Pred_Scaled = budbreak_20_plasticity_aic_pred_scaled,
                              AIC_All_Scaled = budbreak_20_plasticity_aic_all_scaled
                            )

# Order by the best (lowest) AIC from the unscaled models
budbreak_20_plasticity_model_comparison <- budbreak_20_plasticity_model_comparison[order(budbreak_20_plasticity_model_comparison$AIC_Unscaled), ]

budbreak_20_plasticity_model_comparison |> 
  DT::datatable(
    caption = "Bud break plasticity (2020) Model AIC",
    filter="top",
    options = list(pageLength = 11)) |> 
  DT::formatRound(1:3, digits = 2)

```

**Delta AIC**

```{r}
# Calculate AIC for each model
aic_values <- sapply(budbreak_20_plasticity_models_all_scaled, AIC)

# Calculate delta AIC
delta_aic <- aic_values - min(aic_values)

# AIC dataframe
aic_df <- data.frame(
  AIC = aic_values,
  delta_AIC = delta_aic
)
t(aic_df)


# write.csv(t(aic_df), "./tables/budbreak_20_plasticity_AIC.csv")
```

## Diagnostic plots

::: panel-tabset
### Unscaled data

```{r}
par(mfrow = c(2, 2))
plot(budbreak_20_plasticity_models_unscaled[["CMD"]])

plot(budbreak_20_plasticity_models_unscaled[["DD18"]])
```

### Predictors scaled

```{r}
par(mfrow = c(2, 2))
plot(budbreak_20_plasticity_models_pred_scaled[["CMD"]])

plot(budbreak_20_plasticity_models_pred_scaled[["DD18"]])
```

### All variables scaled

```{r}
par(mfrow = c(2, 2))
plot(budbreak_20_plasticity_models_all_scaled[["CMD"]])

plot(budbreak_20_plasticity_models_all_scaled[["DD18"]])
```
:::
:::::

------------------------------------------------------------------------

# Figure 4

```{r echo=T, eval=F}
#| code-fold: true
# image output
png(filename = "./figures/paper_figures/Figure4-1.png", width = 16, height = 14, units = "in", res = 300)

# pdf output
pdf(file = "./figures/paper_figures/Figure4.pdf", width = 16, height = 14)
```

```{r}
#| fig-height: 14
#| fig-width: 16
#| code-fold: true
#| message: false
#| warning: false
# Use layout for more control
layout(matrix(c(1,2,3,4), 2, 2, byrow = TRUE))

# Reduce margins
par(mar = c(2, 1, 2, 2), oma = c(1, 1, 2, 1))

par(cex.lab=2.5) 

# height growth (trait)
plot_in3d(data=data_clim,x="PAS",y="PC2",z="Growth", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=-140,phi=18,
                      xlab="PAS",ylab="Genetic PC2",zlab="Height growth",show_legends=FALSE, color_var = "q2")

# height growth (plasticity)
plot_in3d(data=data_clim,x="PET",y="PC2",z="Growth_plasticity", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=40,phi=18,
                      xlab="PET",ylab="Genetic PC2",zlab="Height growth (p)",show_legends=FALSE, color_var = "q2")


# total dry matter (trait)
plot_in3d(data=data_clim,x="DD18",y="PC2",z="TotalDry", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=140,phi=20,
                      xlab="DD18",ylab="Genetic PC2",zlab="Total dry matter",show_legends=FALSE, color_var = "q2")

# total dry matter (plasticity)
plot_in3d(data=data_clim,x="DD18",y="PC2",z="TotalDry_plasticity", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=140,phi=20,
                      xlab="DD18",ylab="Genetic PC2",zlab="Total dry matter (p)",show_legends=T, color_var = "q2",
                      shape_legend_position="bottomright")

# par(cex.lab=5)
```

```{r echo=T,eval=F}
#| code-fold: true
dev.off()

# Reset par
par(mfrow = c(1, 1))
```

```{r}
#| eval: false
#| echo: false

# testing
ggplot(data_clim, aes(x=PAS, y=PC2, color=Growth)) + 
  geom_point(size=3) + scale_color_gradientn(colors=paletteer_c("ggthemes::Sunset-Sunrise Diverging", 30, direction = 1))

plot_in3d(data=data_clim,x="PET",y="PC2",z="Growth", shape= "Region", 
          pch = c(16,17,15),grid.lines=25, theta=-40,phi=20,
                      xlab="PET",ylab="Genetic PC2",zlab="Height growth", 
          show_legends=FALSE, color_var = "q2")

# old growth plot
# height growth (trait)
plot_in3d(data=data_clim,x="PAS",y="PC2",z="Growth", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=40,phi=20,
                      xlab="PAS",ylab="Genetic PC2",zlab="Height growth",show_legends=FALSE, color_var = "q2")

# better growth plot
plot_in3d(data=data_clim,x="PAS",y="PC2",z="Growth", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=-140,phi=18,
                      xlab="PAS",ylab="Genetic PC2",zlab="Height growth",show_legends=FALSE, color_var = "q2")


plot_in3d(data=data_clim,x="PET",y="PC2",z="Growth_plasticity", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=-40,phi=18,
                      xlab="PET",ylab="Genetic PC2",zlab="Height growth (p)",show_legends=FALSE, color_var = "q2")

```

# Figure S6

```{r echo=T,eval=F}
#| code-fold: true
# image output
png(filename = "./figures/paper_figures/FigureS6-2.png", width = 24, height = 7, units = "in", res = 300)

# pdf output
pdf(file = "./figures/paper_figures/FigureS6.pdf", width = 24, height = 7)
```

```{r}
#| fig-height: 7
#| fig-width: 24
#| code-fold: true
#| message: false
#| warning: false
# Use layout for more control
layout(matrix(c(1,2,3), 1, 3, byrow = TRUE))

# Reduce margins
par(mar = c(2, 1, 2, 2), oma = c(1, 1, 2, 1))

par(cex.lab=2.5) 

# bud set (trait)
plot_in3d(data=data_clim,x="DD_0",y="PC1",z="BudSet_2019", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=40,phi=20,
                      xlab="DD_0",ylab="Genetic PC1",zlab="Bud set (2019)",show_legends=FALSE, color_var = "q2")

# bud break (trait)
plot_in3d(data=data_clim,x="DD18",y="PC1",z="Budbreak_2020", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=110,phi=20,
                      xlab="DD18",ylab="Genetic PC1",zlab="Bud break (2020)",show_legends=FALSE, color_var = "q2")

# bud break (plasticity)
plot_in3d(data=data_clim,x="DD18",y="PC1",z="BudBreak_cGDD_plasticity", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=110,phi=20,
                      xlab="DD18",ylab="Genetic PC1",zlab="Bud break (p)", color_var = "q2",
          show_legends=T, shape_legend_position="bottomright")

# par(cex.lab=5)
```

```{r echo=T,eval=F}
#| code-fold: true
dev.off()

# Reset par
par(mfrow = c(1, 1))
```

```{r}
# better plot for bud break

# bud break (trait)
plot_in3d(data=data_clim,x="DD18",y="PC1",z="Budbreak_2020", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=110,phi=20,
                      xlab="DD18",ylab="Genetic PC1",zlab="Bud break (2020)",show_legends=FALSE, color_var = "q2")

# bud break (plasticity)
plot_in3d(data=data_clim,x="DD18",y="PC1",z="BudBreak_cGDD_plasticity", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=110,phi=20,
                      xlab="DD18",ylab="Genetic PC1",zlab="Bud break (p)", color_var = "q2",
          show_legends=T, shape_legend_position="bottomright")
```

```{r}
#| eval: false
#| echo: false

# old bud break plot

# bud break (trait)
plot_in3d(data=data_clim,x="DD18",y="PC1",z="Budbreak_2020", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=130,phi=20,
                      xlab="DD18",ylab="Genetic PC1",zlab="Bud break (2020)",show_legends=FALSE, color_var = "q2")


# bud break (plasticity)
plot_in3d(data=data_clim,x="DD18",y="PC1",z="BudBreak_cGDD_plasticity", shape= "Region", pch = c(16,17,15),grid.lines=25, theta=130,phi=20,
                      xlab="DD18",ylab="Genetic PC1",zlab="Bud break (p)", color_var = "q2",
          show_legends=T, shape_legend_position="bottomright")
```
