---
title: "Plasticity based on garden PCs"
author: "Anoob Prakash"
date: "02/13/2023"
output:
  rmarkdown::html_document:
    theme: cosmo  
    number_sections: false
    toc: false
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

# About \
**Trait plasticity (as described in the Prakash et al., 2022)**\

To estimate phenotypic plasticity for each trait, we ran LMMs (Model IV, Table S1) that included the fixed effect of garden climate (gPC1, based on each garden’s eigenvector score along PC1 of the climate PCA for the year of measurement), and random effects of Bed (1|Bed) and Family with random intercepts and slopes (gPC1|Family). We used the ‘coef()’ function to extract the random slopes for each Family with gPC1, which provides a measure of plasticity to garden climate (Arnold et al. 2019). To test if the magnitude of plasticity reflected differences due to source climate variability, we modeled the absolute value of plasticity as a function of source climate (sPC1), with a random effect of Population (Model V, Table S1).\

## Data
```{r}
setwd("Z:/Anoob/Polygenic_Adaptation/ANALYSIS")
```

```{r message=FALSE, warning=FALSE}
require(lmerTest)
```



**Source climate**  
```{r}
# setwd("Z:/Anoob/MCMCglmm")

Population_climateNA <- read.csv("../ClimateData/paper1_climate_data/PCA_sprucePops.txt",
                                 header=T, sep="\t")
colnames(Population_climateNA)[colnames(Population_climateNA)=="PC1_sel"] <- "PC1"

Family_climateNA <- read.csv("../ClimateData/paper1_climate_data/PCA_spruceFams.txt",
                                 header=T, sep="\t")
colnames(Family_climateNA)[colnames(Family_climateNA)=="PC1_sel"] <- "PC1"
```

**Garden climate**  
```{r}
Garden_clim <- read.table("../ClimateData/paper1_climate_data/climNA_gardens_2019-20.txt", header=T)
Garden_clim
selVars1 <- c('DD_0','DD18','MAR','PAS','MSP','RH','EXT','CMD','TD','eFFP','PET')
colnames(Garden_clim)[1] <- "Garden"
colnames(Garden_clim)[2] <- "Year"


### estimate the mean climate variables for 2019-2020 across three garden sites
# Subset the data for MD
md_data <- Garden_clim %>% filter(Garden == "MD")
# Calculate column means for numeric columns
md_numeric <- colnames(md_data[-c(1:2)])
md_means <- apply(md_data[, md_numeric], 2, mean)
# Add the means as a new row
md_data_with_means <- rbind(md_data, c(Garden = "MD", Year = "Mean", md_means))

# Subset the data for NC
nc_data <- Garden_clim %>% filter(Garden == "NC")
# Calculate column means for numeric columns
nc_numeric <- colnames(nc_data[-c(1:2)])
nc_means <- apply(nc_data[, nc_numeric], 2, mean)
# Add the means as a new row
nc_data_with_means <- rbind(nc_data, c(Garden = "NC", Year = "Mean", nc_means))

# Subset the data for VT
vt_data <- Garden_clim %>% filter(Garden == "VT")
# Calculate column means for numeric columns
vt_numeric <- colnames(vt_data[-c(1:2)])
vt_means <- apply(vt_data[, vt_numeric], 2, mean)
# Add the means as a new row
vt_data_with_means <- rbind(vt_data, c(Garden = "VT", Year = "Mean", vt_means))

# combined mean data - added to the garden climate data.frame
Garden_clim <- rbind(Garden_clim,md_data_with_means[3,],nc_data_with_means[3,],vt_data_with_means[3,])

# Reset row names and revert columns to numeric
rownames(Garden_clim) <- NULL
Garden_clim[vt_numeric] <- lapply(Garden_clim[vt_numeric], as.numeric)

# Selected cliamte variables
selVars1 <- c('DD18','MAR','PAS','MSP','RH','EXT','CMD','PET')


# create ID for garden
garden_id <- Garden_clim[,1:2]

# selected clim of gardens
data_p1 <- Garden_clim[,selVars1]
```

**Climate PCA**  
```{r}
## Climate PCA
data_pca <- prcomp(data_p1,scale. = T)

# Extract PC axes for plotting
PCAvalues <- data.frame(Garden = Garden_clim$Garden, Year = Garden_clim$Year,
                        data_pca$x)
PCAvalues$Year <- as.factor(PCAvalues$Year)

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(data_pca$rotation), data_pca$rotation)

PCAvalues_mod <- PCAvalues
PCAvalues_mod$Garden[PCAvalues_mod$Garden=="VT"] <- "Vermont"
PCAvalues_mod$Garden[PCAvalues_mod$Garden=="MD"] <- "Maryland"
PCAvalues_mod$Garden[PCAvalues_mod$Garden=="NC"] <- "North_Carolina"
```

**biomass and isotopes**

```{r}
biomass <- read.csv("../TraitData/biomass_2020.csv")
isotope <- read.csv("../TraitData/isotope_2020.csv")

# Region palette
reg_palette <- c("goldenrod2","green2","steelblue")
```

## Plasticity  {.tabset .tabset-fade .tabset-pills}

### Height growth plasticity \

```{r}
growth_2019 <- read.csv("../../MCMCglmm/trait_data/Growth_2019.csv", stringsAsFactors = T)
growth_2020 <- read.csv("../../MCMCglmm/trait_data/Growth_2020.csv", stringsAsFactors = T)

growth_2019$mBed <- paste0(growth_2019$Garden,"_",growth_2019$Bed)
growth_2020$mBed <- paste0(growth_2020$Garden,"_",growth_2020$Bed)

# Create the growth values for 2019-2020
growth_full <- merge(x=subset(growth_2019, select=-c(Height_Fall2019,Growth)), 
                     y=growth_2020[,c("plant_ID","FinalHT")],
                     by.x="plant_ID")
# growth_full <- merge(x=growth_full, 
#                      y=Plasticity[,c("Family","PC1")],
#                      by.x="Family")
growth_full$Growth <- growth_full$FinalHT - growth_full$initial_h


## height growth plasticity
GW <- merge(x=growth_full,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
GW$mBed <- paste0(GW$Garden,"_",GW$Bed)

# plasticity model for bud break 2020
GW_PC1_mod <- lmer(data = GW, na.action = na.omit,
                          Growth ~ PC1  + (1|mBed) + (PC1|Family))

summary(GW_PC1_mod)

# converting the negative values to absolute values to make plasticity values comparable across the traits

gw_PC1 <- coef(GW_PC1_mod)
# hist(testx_PC1$PC1)
gw_PC1 <- gw_PC1$Family
gw_PC1 <- cbind(Family=rownames(gw_PC1),gw_PC1)
rownames(gw_PC1) <- NULL
gw_PC1$abs <- abs(gw_PC1$PC1)
hist(gw_PC1$PC1)
hist(gw_PC1$abs)
gw_PC1 <- gw_PC1[,-2]
colnames(gw_PC1)[3] <- "Growth_plasticity"
gw_PC1 <- gw_PC1[,-2]  
```


### Total dry matter \

```{r}
TDM <- merge(x=biomass,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
TDM$mBed <- paste0(TDM$Garden,"_",TDM$Bed)

#plasticity model
TDM_PC1_mod <- lmer(data = TDM, na.action = na.omit,
                          TotalDry ~ PC1  + (1|mBed) + (PC1|Family))

summary(TDM_PC1_mod)

TDM_PC1 <- ranef(TDM_PC1_mod)
TDM_PC1 <- TDM_PC1$Family
TDM_PC1 <- cbind(Family=rownames(TDM_PC1),TDM_PC1)
rownames(TDM_PC1) <- NULL

TDM_PC1$Family <- as.factor(TDM_PC1$Family)
colnames(TDM_PC1)[colnames(TDM_PC1)=="PC1"] <- "TotalDry_plasticity"
TDM_PC1 <- TDM_PC1[,-2]
head(TDM_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

tdm_PC1 <- coef(TDM_PC1_mod)
# hist(testx_PC1$PC1)
tdm_PC1 <- tdm_PC1$Family
tdm_PC1 <- cbind(Family=rownames(tdm_PC1),tdm_PC1)
rownames(tdm_PC1) <- NULL
tdm_PC1$abs <- abs(tdm_PC1$PC1)
hist(tdm_PC1$PC1)
hist(tdm_PC1$abs)
tdm_PC1 <- tdm_PC1[,-2]
colnames(tdm_PC1)[3] <- "TotalDry_plasticity"
tdm_PC1 <- tdm_PC1[,-2] 
```

### Root:Shoot \
```{r}
RootShoot <- merge(x=biomass,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
RootShoot$mBed <- paste0(RootShoot$Garden,"_",RootShoot$Bed)

#plasticity model
RootShoot_PC1_mod <- lmer(data = RootShoot, na.action = na.omit,
                          RootShoot ~ PC1  + (1|mBed) + (PC1|Family))

summary(RootShoot_PC1_mod)

RootShoot_PC1 <- ranef(RootShoot_PC1_mod)
RootShoot_PC1 <- RootShoot_PC1$Family
RootShoot_PC1 <- cbind(Family=rownames(RootShoot_PC1),RootShoot_PC1)
rownames(RootShoot_PC1) <- NULL

RootShoot_PC1$Family <- as.factor(RootShoot_PC1$Family)
colnames(RootShoot_PC1)[colnames(RootShoot_PC1)=="PC1"] <- "RootShoot_plasticity"
RootShoot_PC1 <- RootShoot_PC1[,-2]
head(RootShoot_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

rootshoot_PC1 <- coef(RootShoot_PC1_mod)
# hist(testx_PC1$PC1)
rootshoot_PC1 <- rootshoot_PC1$Family
rootshoot_PC1 <- cbind(Family=rownames(rootshoot_PC1),rootshoot_PC1)
rownames(rootshoot_PC1) <- NULL
rootshoot_PC1$abs <- abs(rootshoot_PC1$PC1)
hist(rootshoot_PC1$PC1)
hist(rootshoot_PC1$abs)
rootshoot_PC1 <- rootshoot_PC1[,-2]
colnames(rootshoot_PC1)[3] <- "RootShoot_plasticity"
rootshoot_PC1 <- rootshoot_PC1[,-2] 
```


### SideBranch:MainStem \
```{r}
SMR <- merge(x=biomass,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
SMR$mBed <- paste0(SMR$Garden,"_",SMR$Bed)

#plasticity model
SMR_PC1_mod <- lmer(data = SMR, na.action = na.omit,
                          SideBr_MainSt ~ PC1  + (1|mBed) + (PC1|Family))

summary(SMR_PC1_mod)

SMR_PC1 <- ranef(SMR_PC1_mod)
SMR_PC1 <- SMR_PC1$Family
SMR_PC1 <- cbind(Family=rownames(SMR_PC1),SMR_PC1)
rownames(SMR_PC1) <- NULL

SMR_PC1$Family <- as.factor(SMR_PC1$Family)
colnames(SMR_PC1)[colnames(SMR_PC1)=="PC1"] <- "SideBranch_MainStem_plasticity"
SMR_PC1 <- SMR_PC1[,-2]
head(SMR_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

smr_PC1 <- coef(SMR_PC1_mod)
# hist(testx_PC1$PC1)
smr_PC1 <- smr_PC1$Family
smr_PC1 <- cbind(Family=rownames(smr_PC1),smr_PC1)
rownames(smr_PC1) <- NULL
smr_PC1$abs <- abs(smr_PC1$PC1)
hist(smr_PC1$PC1)
hist(smr_PC1$abs)
smr_PC1 <- smr_PC1[,-2]
colnames(smr_PC1)[3] <- "SideBranch_MainStem_plasticity"
smr_PC1 <- smr_PC1[,-2] 
```


### delta13carbon \
```{r}
D13C <- merge(x=isotope,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
D13C$mBed <- paste0(D13C$Garden,"_",D13C$Bed)

#plasticity model
D13C_PC1_mod <- lmer(data = D13C, na.action = na.omit,
                          d13C ~ PC1  + (1|mBed) + (PC1|Family))

summary(D13C_PC1_mod)

D13C_PC1 <- ranef(D13C_PC1_mod)
D13C_PC1 <- D13C_PC1$Family
D13C_PC1 <- cbind(Family=rownames(D13C_PC1),D13C_PC1)
rownames(D13C_PC1) <- NULL

D13C_PC1$Family <- as.factor(D13C_PC1$Family)
colnames(D13C_PC1)[colnames(D13C_PC1)=="PC1"] <- "Delta13Carbon_plasticity"
D13C_PC1 <- D13C_PC1[,-2]
head(D13C_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

d13c_PC1 <- coef(D13C_PC1_mod)
d13c_PC1 <- d13c_PC1$Family
d13c_PC1 <- cbind(Family=rownames(d13c_PC1),d13c_PC1)
rownames(d13c_PC1) <- NULL
d13c_PC1$abs <- abs(d13c_PC1$PC1)
hist(d13c_PC1$PC1)
hist(d13c_PC1$abs)
d13c_PC1 <- d13c_PC1[,-2]
colnames(d13c_PC1)[3] <- "Delta13Carbon_plasticity"
d13c_PC1 <- d13c_PC1[,-2] 
```

### delta15nitrogen \
```{r}
D15N <- merge(x=isotope,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
D15N$mBed <- paste0(D15N$Garden,"_",D15N$Bed)

#plasticity model
D15N_PC1_mod <- lmer(data = D15N, na.action = na.omit,
                          d15N ~ PC1  + (1|mBed) + (PC1|Family))

summary(D15N_PC1_mod)

D15N_PC1 <- ranef(D15N_PC1_mod)
D15N_PC1 <- D15N_PC1$Family
D15N_PC1 <- cbind(Family=rownames(D15N_PC1),D15N_PC1)
rownames(D15N_PC1) <- NULL

D15N_PC1$Family <- as.factor(D15N_PC1$Family)
colnames(D15N_PC1)[colnames(D15N_PC1)=="PC1"] <- "Delta15Nitrogen_plasticity"
D15N_PC1 <- D15N_PC1[,-2]
head(D15N_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

d15n_PC1 <- coef(D15N_PC1_mod)
d15n_PC1 <- d15n_PC1$Family
d15n_PC1 <- cbind(Family=rownames(d15n_PC1),d15n_PC1)
rownames(d15n_PC1) <- NULL
d15n_PC1$abs <- abs(d15n_PC1$PC1)
hist(d15n_PC1$PC1)
hist(d15n_PC1$abs)
d15n_PC1 <- d15n_PC1[,-2]
colnames(d15n_PC1)[3] <- "Delta15Nitrogen_plasticity"
d15n_PC1 <- d15n_PC1[,-2] 
```


### PercentN \
```{r}
PercentN <- merge(x=isotope,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
PercentN$mBed <- paste0(PercentN$Garden,"_",PercentN$Bed)

#plasticity model
PercentN_PC1_mod <- lmer(data = PercentN, na.action = na.omit,
                          percentN ~ PC1  + (1|mBed) + (PC1|Family))

summary(PercentN_PC1_mod)

PercentN_PC1 <- ranef(PercentN_PC1_mod)
PercentN_PC1 <- PercentN_PC1$Family
PercentN_PC1 <- cbind(Family=rownames(PercentN_PC1),PercentN_PC1)
rownames(PercentN_PC1) <- NULL

PercentN_PC1$Family <- as.factor(PercentN_PC1$Family)
colnames(PercentN_PC1)[colnames(PercentN_PC1)=="PC1"] <- "PercentN_plasticity"
PercentN_PC1 <- PercentN_PC1[,-2]
head(PercentN_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

percentN_PC1 <- coef(PercentN_PC1_mod)
percentN_PC1 <- percentN_PC1$Family
percentN_PC1 <- cbind(Family=rownames(percentN_PC1),percentN_PC1)
rownames(percentN_PC1) <- NULL
percentN_PC1$abs <- abs(percentN_PC1$PC1)
hist(percentN_PC1$PC1)
hist(percentN_PC1$abs)
percentN_PC1 <- percentN_PC1[,-2]
colnames(percentN_PC1)[3] <- "PercentN_plasticity"
percentN_PC1 <- percentN_PC1[,-2] 
```

### CN ratio \

```{r}
CNR <- merge(x=isotope,y=PCAvalues_mod[PCAvalues_mod$Year=="Mean",], all.x=TRUE)
CNR$mBed <- paste0(CNR$Garden,"_",CNR$Bed)

#plasticity model
CNR_PC1_mod <- lmer(data = CNR, na.action = na.omit,
                          CN_ratio ~ PC1  + (1|mBed) + (PC1|Family))

summary(CNR_PC1_mod)

CNR_PC1 <- ranef(CNR_PC1_mod)
CNR_PC1 <- CNR_PC1$Family
CNR_PC1 <- cbind(Family=rownames(CNR_PC1),CNR_PC1)
rownames(CNR_PC1) <- NULL

CNR_PC1$Family <- as.factor(CNR_PC1$Family)
colnames(CNR_PC1)[colnames(CNR_PC1)=="PC1"] <- "CN_ratio_plasticity"
CNR_PC1 <- CNR_PC1[,-2]
head(CNR_PC1) 

# converting the negative values to absolute values to make plasticity values comparable across the traits

cnr_PC1 <- coef(CNR_PC1_mod)
cnr_PC1 <- cnr_PC1$Family
cnr_PC1 <- cbind(Family=rownames(cnr_PC1),cnr_PC1)
rownames(cnr_PC1) <- NULL
cnr_PC1$abs <- abs(cnr_PC1$PC1)
hist(cnr_PC1$PC1)
hist(cnr_PC1$abs)
cnr_PC1 <- cnr_PC1[,-2]
colnames(cnr_PC1)[3] <- "CN_ratio_plasticity"
cnr_PC1 <- cnr_PC1[,-2] 
```

## combine data\

```{r}
require(tidyverse)
# combine the data frame for biomass and isotope blups
plasticity <- list(as.data.frame(gw_PC1),as.data.frame(tdm_PC1),as.data.frame(rootshoot_PC1), 
                   as.data.frame(smr_PC1), # biomass
                   as.data.frame(d13c_PC1),as.data.frame(d15n_PC1), as.data.frame(percentN_PC1),as.data.frame(cnr_PC1)) 
#merge all data frames in list
plasticity <- plasticity %>% reduce(full_join, by='Family')

# write.csv(plasticity,"../TraitData/biomass_isotope_plasticity_2020.csv")
```



***

# References\

*Arnold PA, Nicotra AB, Kruuk LEB. 2019 Sparse evidence for selection on phenotypic plasticity in response to temperature. Phil. Trans. R. Soc. B 374, 20180185.* \

*Prakash, A., DeYoung, S., Lachmuth, S., Adams, J. L., Johnsen, K., Butnor, J. R., ... & Keller, S. R. (2022). Genotypic variation and plasticity in climate-adaptive traits after range expansion and fragmentation of red spruce (Picea rubens Sarg.). Philosophical Transactions of the Royal Society B, 377(1848), 20210008.*\

