---
title: "Qpc - September 2025"
author: Anoob Prakash
lightbox: true
format: 
  html:
    code-overflow: wrap
    toc: true
    embed-resources: true
    code-tools: true
    code-fold: false
knitr:
  opts_knit:
    root.dir: "Z:/Anoob/Polygenic_Adaptation/ANALYSIS"
editor_options: 
  chunk_output_type: console
---

# Packages required

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# packages
require(stringr)
require(viridis)
require(lmerTest)
require(tidyverse)
require(qvalue)
require(readxl)
require(RcppCNPy)
require(reticulate)
require(pheatmap)
require(tidyr)
require(reshape2)
require(plyr)
require(cowplot)
require(gridExtra)

# function for Qpc
source("./Qpc_function.R")
```

# Data

```{r}
# Meta data
Meta <- read.table("../../MCMCglmm/Exome/RS_Exome_metadata.txt", sep="\t",header=T)
bluptable <- read.csv("../TraitData/trait_blups_complete.csv")

## Covariance between inds 
### Eigen 2
cov <- as.matrix(read.table("filtering_the_snps_across_REGION/selection/Full_Sampling_minMAF_filtering_run2_0.1_PCA_eigen2_retest_July2025.cov"))

dim(cov)
```

```{r}
## Generating meta data for downstream analysis
# 324 fam
# create bam list for the traits being studied
labels <- read.table("black_spruce/new_324RSfam_bam_VACC.list", stringsAsFactors = F)
labels$V1 <- str_replace_all(labels$V1,"/netfiles02/kellrlab/datashare/Spruce/exome_capture/WES_mapping/Mapping_WES/ref_Pmariana/BWA/Prubens/NSF/","")
labels$V1 <- str_replace_all(labels$V1,".sorted.rmdup.bam","")
colnames(labels) <- "Family"

labels <- as.data.frame(labels) %>%
              left_join(Meta[, c("Pop", "Family", "Region")], by = c("Family"))

fam_label <- labels$Family
```

## PCA - pcangsd covariance matrix

```{r}
eig <- eigen(cov)

# Extract the eigenvalues
eigenvalues <- eig$values

# Calculate the proportion of variance explained
var_explained <- eigenvalues / sum(eigenvalues)
var_explained[1:5]

# add the meta data to the PC components
pca_scores <- as.data.frame(eig$vectors)
colnames(pca_scores) <- paste0("PC", 1:ncol(pca_scores))

pca_scores <- cbind(as.data.frame(labels),pca_scores)
colnames(pca_scores)[1] <- "Family"

# pca_scores <- merge(x=Meta[,c(1,3,5)], y=pca_scores, all.y=T, by="Family")
pca_scores$Region <- factor(pca_scores$Region, levels=c("C","M","E"),
                                               labels=c("Core","Margin","Edge"))

# plot(pca_scores$PC1,pca_scores$PC2)

PCplot <- ggplot(pca_scores[!is.na(pca_scores$Region),], 
                 aes(x = PC1, y = PC2, text=Family, color=Region, shape=Region)) +
                geom_point(aes(color=Region),  pch = 20,alpha=0.7, size=5) + 
                labs(x="PC1 (2.5%)", y="PC2 (2.1%)") +
          scale_color_manual(values = c("Core" = "goldenrod2","Margin" = "green2","Edge" = "steelblue")) +
          theme_minimal() +
          theme(
          axis.text.x=element_text(size=14, angle=0, vjust=0.3),
          axis.text.y=element_text(size=14),
          axis.title =element_text(size=14, face = "bold"),
          plot.title=element_text(size=11),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "top",
          panel.ontop = TRUE,
          panel.background = element_rect(fill = "transparent"),
          panel.grid = element_line(color="black"))  

PCplot

# ggsave("./figures/PCA_red_spruce_cov.png", width = 2000, height = 1200, units = "px", type = "cairo")
# ggsave("./figures/PCA_red_spruce_cov.pdf", width = 2000, height = 1200, units = "px")
```

### Make Kinship Matrix

```{r eval=F}
## Make K matrix
# run on Wright (HPC cluster)

# Dosage file
require(RcppCNPy)
require(reticulate)


### eigen 2
# Dosage file
require(RcppCNPy)
require(reticulate)

## Wright 
setwd("/home/Anoob/mydata/Anoob/Polygenic_Adaptation/ANALYSIS/filtering_the_snps_across_REGION/selection/")

np <- import("numpy")
data <- as.data.frame(t(np$load("Full_Sampling_minMAF_filtering_run2_0.1_PCA_eigen2_retest_July2025.dosage.npy"))) ## version min maf 0.1
dim(data) # 742,006

# convert dosage to AF
Afreqs = data/2

# Now, randomly subsample 100K loci
total_sites <- dim(Afreqs)[2]
Loci = seq(1,total_sites,1) 
Rand.loci = sample(Loci, 1e5, replace=F)
G = Afreqs[,Rand.loci]
dim(G)

# modified function  -- working final
scaleFactor = sqrt(mean(colMeans(G) * (1 - colMeans(G))))
    myM = dim(G)[1]
    myT = matrix(data = -1/myM, nrow = myM - 1, ncol = myM)
    diag(myT) = (myM - 1)/myM
    myGstand = (myT %*% as.matrix(G))/scaleFactor
    myK = cov(t(myGstand))
    
# kinship used in the paper
saveRDS(myK,"/home/Anoob/mydata/Anoob/Polygenic_Adaptation/ANALYSIS/filtering_the_snps_across_REGION/kinship_matrix/myK_from_modified_make_KE_run2_0.1_dosage_1e5_eigen2_retest_July2025")

# original function
make_k_E <- function (myG)
{
  scaleFactor = sqrt(mean(colMeans(myG) * (1 - colMeans(myG))))
    myM = dim(myG)[1]
    myT = matrix(data = -1/myM, nrow = myM - 1, ncol = myM)
    diag(myT) = (myM - 1)/myM
    myGstand = (myT %*% myG)/scaleFactor
    myK = cov(t(myGstand))
    return(myK)
}

myK <- make_k_E(G)
# end of original function

```

# Q~PC~ Analysis

:::panel-tabset
## Heatmap

```{r}

myK <- readRDS("filtering_the_snps_across_REGION/kinship_matrix/myK_from_modified_make_KE_run2_0.1_dosage_1e5_eigen2_retest_July2025")

# bam list
kinship_labels <- read.table("black_spruce/new_324RSfam_bam_VACC.list", stringsAsFactors = F)
kinship_labels$V1 <- str_replace_all(kinship_labels$V1,"/netfiles02/kellrlab/datashare/Spruce/exome_capture/WES_mapping/Mapping_WES/ref_Pmariana/BWA/Prubens/NSF/","")
kinship_labels$V1 <- str_replace_all(kinship_labels$V1,".sorted.rmdup.bam","")
kinship_labels <- kinship_labels$V1

# plot the kinship matrix map
heat_K <- myK

row.names(heat_K) = kinship_labels[1:dim(heat_K)[1]]
colnames(heat_K) <- kinship_labels[1:dim(heat_K)[1]]

# my_sample_col <- data.frame(1:332,1:332)
my_sample_col <- as.data.frame(colnames(heat_K))
colnames(my_sample_col) <- "Family"

my_sample_col <- left_join(my_sample_col,Meta,by="Family")
my_sample_col$Species <- NULL

my_sample_col$Region <- factor(my_sample_col$Region,levels=c("C","M","E"),
                               labels=c("Core","Margin","Edge"))
Region <- as.data.frame(as.character(my_sample_col$Region))
colnames(Region) <- "Region"
rownames(Region) <- colnames(heat_K)

# Specify colors
ann_colors = list(
    # Species = c(BlackSpruce="black", RedSpruce="firebrick"),
    Region = c(Core = "goldenrod2", Margin = "green2", 
               Edge = "steelblue")
)


diag(heat_K) <- NA

# png(file="./figures/Qpc/heatmap_pcangsd_cov.png", width=3000,height=2800,units = "px")
heatmap <- pheatmap(heat_K, 
           annotation_row = Region,
           annotation_col = Region,
           annotation_colors = ann_colors,
           # cutree_rows = 2,
           # cutree_cols = 2,
           col=viridis(400),
           show_rownames = F,
           annotation_names_row = F,
           show_colnames = F,
           labels_row = F)

heatmap

# ggsave("./figures/Qpc/heatmap_pcangsd_cov.png", width = 3000, height = 2800, units = "px", type = "cairo")
```

## PCA of K-matrix

```{r}
myK <- readRDS("filtering_the_snps_across_REGION/kinship_matrix/myK_from_modified_make_KE_run2_0.1_dosage_1e5_eigen2_retest_July2025")
#do the eigen decomposition
eigF = eigen(myK)
myU = eigF$vectors
myLambdas = eigF$values

# calculate the PC cutoffs we're using for tests of selection
varexp = myLambdas/sum(myLambdas)
sumexp = sapply(1:length(varexp), function(x){sum(varexp[1:x])})

plot(sumexp[1:5])

sumexp[2] # take the first 4 PCs to account for the population structure 
cutoff <- sumexp[2]
#get cutoffs for how many pcs to look at
pcmax = which(sumexp > (cutoff-0.005))[1]
pcmax

# Extract the eigenvalues
eigenvalues <- eigF$values

# Calculate the proportion of variance explained
var_explained <- eigenvalues / sum(eigenvalues)
var_explained[1:5]


# add the meta data to the PC components
pca_scores <- as.data.frame(eigF$vectors)
colnames(pca_scores) <- paste0("PC", 1:ncol(pca_scores))

# add row names
pca_scores <- cbind(as.data.frame(kinship_labels[1:323]),pca_scores)
colnames(pca_scores)[1] <- "Family"

# add meta data
pca_scores <- left_join(x=pca_scores, y=Meta, by="Family")

# pca_scores <- merge(x=Meta[,c(1,3,5)], y=pca_scores, all.y=T, by="Family")
pca_scores$Region <- factor(pca_scores$Region, levels=c("C","M","E"),
                                               labels=c("Core","Margin","Edge"))

par(mfrow=c(1,1))
# varinace explained
plot(myLambdas, bty="n", xlab = "PCs", ylab = "Eigenvalues")
abline(v = pcmax, col = viridis(6)[3], lwd=2)
tailCutoff=0.9

 mydfs <- apply(bluptable[,-1], 2, function(x){calcQpc(
                                                myZ = x, 
                                                myU = eigF$vectors, 
                                                myLambdas = eigF$values, 
                                                myPCcutoff=cutoff,
                                                vapcs = 110,    # the number of pcs used to estimate Va
                                                tailCutoff = 0.9 
                                              )})
```

**Variance explained by PC axis**

```{r}
var_explained[1:5]*100
```

```{r}
PCplot <- ggplot(pca_scores[!is.na(pca_scores$Region),], aes(x = PC1, y = PC2, text=Family, color=Region, shape=Region)) +
                geom_point(aes(color=Region),  shape=1, alpha=0.9, size=5) + 
                labs(x="PC1 (7.7%)", y="PC2 (6.1%)") +
          scale_color_manual(values = c("Core" = "goldenrod2","Margin" = "green2","Edge" = "steelblue")) +
          theme_minimal() +
          theme(
          axis.text.x=element_text(size=14, angle=0, vjust=0.3),
          axis.text.y=element_text(size=14),
          axis.title =element_text(size=14, face = "bold"),
          plot.title=element_text(size=11),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "top",
          panel.ontop = TRUE,
          panel.background = element_rect(fill = "transparent"),
          panel.grid = element_line(color="black"))  

PCplot

# ggsave("./figures/Qpc/make_K_E_PCA1_PCA2.png", width = 2000, height = 2000, units = "px", type = "cairo")

PCplot2 <- ggplot(pca_scores[!is.na(pca_scores$Region),], aes(x = PC1, y = PC3, text=Family, color=Region, shape=Region)) +
                geom_point(aes(color=Region),  shape=1, alpha=0.9, size=5) + 
                labs(x="PC1 (7.7%)", y="PC3 (0.4%)") +
          scale_color_manual(values = c("Core" = "goldenrod2","Margin" = "green2","Edge" = "steelblue")) +
          theme_minimal() +
          theme(
          axis.text.x=element_text(size=14, angle=0, vjust=0.3),
          axis.text.y=element_text(size=14),
          axis.title =element_text(size=14, face = "bold"),
          plot.title=element_text(size=11),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          legend.position = "top",
          panel.ontop = TRUE,
          panel.background = element_rect(fill = "transparent"),
          panel.grid = element_line(color="black")) 

PCplot2

# ggsave("./figures/Qpc/make_K_E_PCA2_PCA3.png", width = 2000, height = 2000, units = "px", type = "cairo")
```

## Generate pvalues and qvalues

```{r}
getqvalues <- function(ptable){
  qobj = qvalue(p = c(ptable))
  myqvals = matrix(qobj$qvalues, nrow=dim(ptable)[1])
  return(myqvals)
}

# No. of traits observed - 1
traitObs <- 21

allpvals = sapply(1:22, function(x){mydfs[[x]]$pvals})
myqvals = getqvalues(allpvals)
```

**plot the pvalues**

-   before FDR correction\

```{r}
#| fig-height: 8
#| fig-width: 6
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,13,1,1), xpd=TRUE, mfrow=c(1,1)) # par(mar=c(8,14,2,2),
mycol = c(viridis(6, direction=1)[1:4], "white")
image(allpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)  
legend(-0.5,-0.20, levels(mysig2), fill=mycol, bty="n", horiz=T)
```

**plot the qvalues**

-   after FDR correction\

```{r}
#| fig-height: 8
#| fig-width: 6
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,13,1,1), xpd=TRUE, mfrow=c(1,1)) # par(mar=c(8,14,2,2),
mycol = c(viridis(6, direction=1)[1:4], "white")
image(myqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)  
legend(-0.5,-0.20, levels(mysig2), fill=mycol, bty="n", horiz=T)
```

## Sensitivity testing

**Testing the sensitivity for the choice of PCs used to estimate Va**

-   Use all PCs\
-   Use 110 PCs\
-   Use 50 PCs\

```{r}
trait_num <- 22
## try with all PCS below the cutoff 
mydfsAllpcs = apply(bluptable[,-1], 2, function(x){calcQpc(
  myZ = x, myU = eigF$vectors, myLambdas = eigF$values, myPCcutoff=cutoff, vapcs = 323, tailCutoff = 1
)})

allpvalsAllpcs = sapply(1:trait_num, function(x){mydfsAllpcs[[x]]$pvals})
myqvalsAllpcs = getqvalues(allpvalsAllpcs)

#use only 50 PCs
mydfs50pcs = apply(bluptable[,-1], 2, function(x){calcQpc(
  myZ = x, myU = eigF$vectors, myLambdas = eigF$values, myPCcutoff=cutoff, vapcs = 50, tailCutoff = 0.9
)})

allpvals50pcs = sapply(1:trait_num, function(x){mydfs50pcs[[x]]$pvals})
myqvals50pcs = getqvalues(allpvals50pcs)

```

**changing the number of PCs used - before FDR**

```{r}
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,10,2,2), xpd=TRUE, mfrow=c(1,1), mfrow=c(1,3))
mycol = c(viridis(6, direction=1)[1:4], "white")

## changing the number of PCs used - before FDR 
image(allpvalsAllpcs, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(allpvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(allpvals50pcs, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

par(mfrow=c(1,1))
```

**changing the number of PCs used - after FDR**

```{r}
layout(matrix(1, nrow=1, ncol=1))
mysig2 =  cut((1:1000/1000), c(0,0.001,0.01,0.05,0.1,1)) #for legend
par(mar=c(8,10,2,2), xpd=TRUE, mfrow=c(1,1), mfrow=c(1,3))
mycol = c(viridis(6, direction=1)[1:4], "white")

## changing the number of PCs used - after FDR
image(myqvalsAllpcs, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(myqvals, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

image(myqvals50pcs, col=mycol, xaxt="n", yaxt="n", bty="n", breaks=c(0,0.001,0.01,0.05,0.1,1))
axis(1, at=seq(0,1, length=pcmax), las=2, label=1:pcmax)
axis(2, at=(0:traitObs)/traitObs, labels = colnames(bluptable[,-1]), las=2)
legend(-0.2,-0.15, levels(mysig2), fill=mycol, bty="n", horiz=T)

par(mfrow=c(1,1))
```
::: 

# Q~PC~ visualization  
:::panel-tabset  

## Data wrangling
```{r}
### pvalues
mypvals_df <- as.data.frame(allpvals)
colnames(mypvals_df) <- c("Budbreak_2020","BudSet_2019","BudSet_2020","Height_growth",
                          "BudBreak_2020_plasticity","BudSet_2019_plasticity",
                          "BudSet_2020_plasticity","Growth_plasticity",
                          "TotalDry_plasticity","RootShoot_plasticity",
                          "SideBranch_MainStem_plasticity","Delta13Carbon_plasticity",
                          "Delta15Nitrogen_plasticity","PercentN_plasticity",
                          "CN_ratio_plasticity","TotalDry_biomass","RootShoot","SideBranch_MainStem",
                          "Delta13Carbon","Delta15Nitrogen","PercentN","CN_ratio")
mypvals_df$PC <- seq(1:nrow(mypvals_df))

# mypvals_long <- melt(mypvals_df)

mypvals_df <- mypvals_df %>% pivot_longer(names_to = "Trait", cols = Budbreak_2020:CN_ratio)

mypvals_df$Sig <- cut(mypvals_df$value, c(0,0.001,0.01,0.05,0.1,1))
colnames(mypvals_df)[3] <- "pvalue"

mypvals_df2 <- mypvals_df
mypvals_df2$Category <- NA
mypvals_df2$mergeColumn <- paste0(mypvals_df2$Trait,"_",mypvals_df2$PC)

### qvalues
myqvals_df <- as.data.frame(myqvals)
colnames(myqvals_df) <- c("Budbreak_2020","BudSet_2019","BudSet_2020","Height_growth",
                          "BudBreak_2020_plasticity","BudSet_2019_plasticity",
                          "BudSet_2020_plasticity","Growth_plasticity",
                          "TotalDry_plasticity","RootShoot_plasticity",
                          "SideBranch_MainStem_plasticity","Delta13Carbon_plasticity",
                          "Delta15Nitrogen_plasticity","PercentN_plasticity",
                          "CN_ratio_plasticity","TotalDry_biomass","RootShoot","SideBranch_MainStem",
                          "Delta13Carbon","Delta15Nitrogen","PercentN","CN_ratio")
myqvals_df$PC <- seq(1:nrow(myqvals_df))

# myqvals_long <- melt(myqvals_df)

myqvals_df <- myqvals_df %>% pivot_longer(names_to = "Trait", cols = Budbreak_2020:CN_ratio)

colnames(myqvals_df)[3] <- "qvalue"

# set FDR cutoff
myqvals_df$FDR <- ifelse(myqvals_df$qvalue<=0.1,myqvals_df$qvalue,NA)

myqvals_df2 <- myqvals_df
myqvals_df2$mergeColumn <- paste0(myqvals_df2$Trait,"_",myqvals_df2$PC)

## Merge pvals and qvals
pq_vals <- merge(mypvals_df2,myqvals_df2)

par(mfrow=c(2,1))
hist(pq_vals$pvalue)
hist(pq_vals$qvalue)
par(mfrow=c(1,1))
```

**Create categories for the facet wrapping**

```{r}
pq_vals <- pq_vals %>%
                  mutate(Category = case_when(
                    endsWith(Trait, "plasticity") ~ "Trait Plasticity",
                    endsWith(Trait, "2019") ~ "Phenology",
                    endsWith(Trait, "2020") ~ "Phenology",
                    endsWith(Trait, "Height_growth") ~"Growth & Biomass",
                    endsWith(Trait,"Stem") ~ "Growth & Biomass",
                    endsWith(Trait, "TotalDry_biomass") ~ "Growth & Biomass",
                    endsWith(Trait, "RootShoot") ~ "Growth & Biomass",
                    endsWith(Trait,"Nitrogen") ~ "Physiology",
                    endsWith(Trait,"Carbon") ~ "Physiology",
                    endsWith(Trait, "PercentN") ~ "Physiology",
                    endsWith(Trait,"CN_ratio") ~ "Physiology"
                    ))
pq_vals$Trait <- factor(pq_vals$Trait,levels = c("SideBranch_MainStem","RootShoot","TotalDry_biomass",
                                                 "Height_growth",
                                                 "CN_ratio","PercentN","Delta15Nitrogen","Delta13Carbon",
                                                 "Growth_plasticity","TotalDry_plasticity","RootShoot_plasticity",
                                                 "SideBranch_MainStem_plasticity","Delta13Carbon_plasticity",
                                                 "Delta15Nitrogen_plasticity","PercentN_plasticity",
                                                 "CN_ratio_plasticity","BudBreak_2020_plasticity",
                                                 "BudSet_2020_plasticity","BudSet_2019_plasticity",
                                                 "Budbreak_2020","BudSet_2020","BudSet_2019"))
# rename the category labels for the manuscript
pq_vals$Figure_axis_text <- revalue(pq_vals$Trait,
                                               c(SideBranch_MainStem = "Side branch:Main stem",
                                               RootShoot = "Root:Shoot",
                                               TotalDry_biomass = "Total dry biomass",
                                               Height_growth = "Height growth",
                                               CN_ratio = "C:N",
                                               PercentN = "Percent N",
                                               Delta15Nitrogen = "delta 15N",
                                               Delta13Carbon = "delta 13C",
                                               Growth_plasticity = "Height growth",
                                               TotalDry_plasticity = "Total dry biomass",
                                               RootShoot_plasticity = "Root:Shoot",
                                               SideBranch_MainStem_plasticity = "Side branch:Main stem",
                                               Delta13Carbon_plasticity = "delta 13C",
                                               Delta15Nitrogen_plasticity = "delta 15N",
                                               PercentN_plasticity = "Percent N",
                                               CN_ratio_plasticity = "C:N",
                                               BudBreak_2020_plasticity = "Bud break (2020)",
                                               BudSet_2020_plasticity = "Bud set (2020)",
                                               BudSet_2019_plasticity = "Bud set (2019)",
                                               Budbreak_2020 = "Bud break (2020)",
                                               BudSet_2020 = "Bud set (2020)",
                                               BudSet_2019 = "Bud set (2019)"))




# save eigen2 qvals
eigen2_qvals <- pq_vals
```

## Q~PC~ plot   

- with FDR<0.1 cutoff  

```{r}
#| warning: false
#| fig-height: 4
#| fig-width: 14
# pdf dim = 6.65w, 10.86h in inches
# plotting Qpc plot - in USE and improved

require(paletteer)
sig_levels <- as.character(levels(pq_vals$Sig))

pq_vals$Figure_axis_text <- factor(pq_vals$Figure_axis_text,
                                   levels=c("Height growth","Total dry biomass","Root:Shoot",
                                   "Side branch:Main stem","Bud set (2019)","Bud set (2020)",
                                   "Bud break (2020)","delta 13C","delta 15N","Percent N","C:N"))

pq_vals_modified <- pq_vals[pq_vals$PC==c(1,2),]


# Horizontal Qpc plot
qpc_plot <- ggplot(pq_vals, aes(x = Figure_axis_text, y = PC)) + 
                    geom_raster(aes(fill=Sig),hjust =0.5,
                                vjust =0.5) + 
                    scale_fill_manual(values = c("#586085", "#d65b5a", "#ea9656","#eeeee4")) +
                    # highlight qvalue FDR <=0.05
                    geom_point(aes(size=ifelse(FDR, "dot", "no_dot"),color =ifelse(FDR<=0.1, 'white'))) +  
                       scale_color_identity() +
                       scale_size_manual(values=c(dot=4, no_dot=NA), guide="none") +   
                    # coord_cartesian(xlim=c(0.75, 5)) +
                    coord_cartesian(ylim=c(0.65, 2.35)) +
                    scale_y_continuous(breaks = 1:2, labels = as.character(1:2)) +
                    scale_x_discrete(expand = c(0, 0)) +
                    labs(x="", y="PC", title="",fill="Significance level") +
                    theme_bw() + 
                    theme(axis.text.x=element_text(size=10, angle=45, hjust=1),
                                       axis.text.y=element_text(size=14),
                                       plot.title=element_text(size=12),
                                       strip.text = element_text(size = 15),
                                       legend.spacing.x = unit(2.0, 'cm'),
                                       # plot.margin = margin(t=0, # top margin
                                       #                      r=0.25, # right margin
                                       #                      b=0, # bottom margin
                                       #                      l=0.5, # left margin
                                       #                      unit="cm"),
                                       panel.grid.major = element_blank(),
                                       panel.grid.minor = element_blank(),
                                       panel.ontop = TRUE,
                                       panel.background = element_rect(fill = "transparent"),
                                       panel.grid = element_line(color="black"),
                                       legend.position = "top") + 
                    facet_grid(.~Category,scales = "free", space = "free") 
qpc_plot 

# ggsave("./figures/Qpc/Qpc_BS_ref_wide.png", width = 3000, height = 1200, units = "px", type = "cairo")
# 
# ggsave("./figures/Qpc/Qpc_BS_ref_horizontal_paper_legend_modified.png", width = 3000, height = 1200,
#        units = "px", type = "cairo")
```

## Incorporate the ancestry score

```{r}
# Meta data
Meta <- read.table("../../BlackSpruce/Info_samples_all.txt", sep="\t",header=T)

# Get ID and pop info for each individual
pop<-read.table("filtering_the_snps_across_REGION/REGIONS/324RSfam_18BSfam_bam_VACC.list")
pop$Family <- str_replace_all(pop$V1,"/netfiles02/kellrlab/datashare/Spruce/exome_capture/WES_mapping/Mapping_WES/ref_Pmariana/BWA/Prubens/NSF/","")
pop$Family <- str_replace_all(pop$Family,"/netfiles02/kellrlab/datashare/Spruce/exome_capture/WES_mapping/Mapping_WES/ref_Pmariana/BWA/Pmariana/","")

pop$Family <- str_replace_all(pop$Family,".sorted.rmdup.bam","")
pop <- pop$Family

# row_meta <- dplyr::inner_join(Meta,pop,by="Family")
# row_meta$row_index <- as.numeric(row.names(row_meta))
# 
# pop <- row_meta[,colnames(row_meta)!=c("V1","row_index")]


# Read inferred admixture proportions file
q<-read.table("filtering_the_snps_across_REGION/admix/Full_Sampling_intersect_poly_updated_with_BS_samples_PCA_admix_K2.admix.2.Q")
colnames(q) <- c("q1","q2")

admix.id = as.data.frame(cbind(pop, q))
names(admix.id) <- c("Family","q1","q2")
# write.csv(admix.id,"data/admix.csv",row.names = F)
# admix.id = admix.id[,c("Family","q1","q2")] # q2 is the proportion of BS ancestry in the sample

#subset out the trait data ancestry scores from pcangsd admix values  
ancestry <- admix.id[admix.id$Family %in% kinship_labels,]
```

```{r class.source = 'fold-hide'}
#read in population data and merge with traits and eigenvectors
fgmerge <- left_join(pca_scores,bluptable,by="Family")
fgmerge$Region <- as.factor(fgmerge$Region)
# fgmerge$Region <- plyr::revalue(fgmerge$Region, c("C"="Core", "M"="Margin", "E" = "Edge")) 
fgmerge$Region <- factor(fgmerge$Region,levels = c("Core", "Margin", "Edge"))

palette(viridis(3))
nicepops = c('Core','Margin','Edge')
indvar = sqrt(diag(as.matrix(myK)))
```

```{r}
#| warning: false

fgmerge <- merge(x=fgmerge,y=ancestry,all.x=T,by="Family")

fgmerge <- fgmerge[,c("Family","PC1","PC2","PC3","PC4","Pop","Tree","State","Region","Date","Latitude","Longitude","Elevation",
                      "DBH","Budbreak_2020","BudSet_2019","BudSet_2020","Growth","BudBreak_cGDD_plasticity",
                      "BudSet_2019_plasticity","BudSet_2020_plasticity","Growth_plasticity","TotalDry_plasticity",
                      "RootShoot_plasticity","SideBranch_MainStem_plasticity","Delta13Carbon_plasticity",
                      "Delta15Nitrogen_plasticity","PercentN_plasticity","CN_ratio_plasticity","TotalDry","RootShoot",
                      "SideBranch_MainStem","delta13C","delta15N","percentN","CN_ratio","q1","q2")]

# write out the results 
# write.csv(fgmerge,"data/Qpc_data.csv",row.names = F)
```


```{r}
# PCA - variance explained
var_explained[1:5]*100
# PC plots - PC plot for paper
PC1_PC2_plot <- ggplot(fgmerge, aes(x = PC1, y = PC2,color = q2,shape=Region) ) +
                        geom_point( size=4, alpha=0.9) +
                labs(x="PC1 (7.7%)", y="PC2 (6.1%)") +
                # scale_shape_manual() +
                # scale_color_manual(values=c("goldenrod2", "green2", "steelblue")) +
                scale_color_gradient("Ancestry",low = "grey95",high = "#251714FF") +
                # geom_abline(intercept=mean(fgmerge[[y]]),slope=1.96*CI[1],lty=2, col=viridis(6)[3], lwd=1) +
                # geom_abline(intercept=mean(fgmerge[[y]]),slope=-1.96*CI[1],lty=2, col=viridis(6)[3], lwd=1) +
                # geom_smooth(aes(x=PC2, y=Growth),method="lm",se=FALSE,formula = y ~ x,inherit.aes = FALSE, lwd=1) +
                theme_bw(base_size = 11, base_family = "Times") +
                theme(axis.text=element_text(size=14), 
                      axis.title=element_text(size=12),
                      # panel.background = element_blank(), 
                      legend.background = element_blank(), 
                      panel.grid = element_blank(), 
                      panel.background = element_rect(fill = "transparent"), 
                      legend.text=element_text(size=rel(.8)), 
                      strip.text = element_text(size=15),
                      legend.position = "none")
PC1_PC2_plot
#FF9898FF #D9636CFF #A91E45FF #691238FF #251714FF

# ggsave("./figures/Qpc/make_K_E_PCA1_PCA2_ancestry_coloring.png", width = 2500, height = 2000, units = "px", type = "cairo")

var_explained[1:5]
# PC plots - PC plot for paper
PC_region_plot <- ggplot(fgmerge, aes(x = PC1, y = PC2,text=Family, color=Region, shape=Region)) +
                  geom_point(aes(color=Region),  shape=1, alpha=0.9, size=5) + 
                  labs(x="PC1 (8%)", y="PC2 (0.6%)") +
                  scale_color_manual(values = c("Core" = "goldenrod2",
                                                "Margin" = "green2",
                                                "Edge" = "steelblue")) +
                  theme_minimal() +
                  theme(
                  axis.text.x=element_text(size=14, angle=0, vjust=0.3),
                  axis.text.y=element_text(size=14),
                  axis.title =element_text(size=14, face = "bold"),
                  plot.title=element_text(size=11),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  legend.position = "top",
                  panel.ontop = TRUE,
                  panel.background = element_rect(fill = "transparent"),
                  panel.grid = element_line(color="black"))



# ggsave("./figures/Qpc/make_K_E_PCA1_PCA2.png", width = 2000, height = 2000, units = "px", type = "cairo")



```

## Choice of Eigen

```{r}
cov_20 <- as.matrix(read.table("filtering_the_snps_across_REGION/selection/Full_Sampling_minMAF_filtering_run2_0.1_PCA_eigen20_retest_July2025.cov"))

dim(cov_20)

eig_20 <- eigen(cov_20)
# plot(eig$vectors[,1:2], xlab="PC1", ylab="PC2", main="PCAngsd")

# Extract the eigenvalues
eigenvalues_20 <- eig_20$values

# Calculate the proportion of variance explained
var_explained_20 <- eigenvalues_20 / sum(eigenvalues_20)
barplot(var_explained_20[1:25])

# Suppose var_explained_20 is your vector of proportions for the first 25 PCs
df <- data.frame(
  PC = factor(1:25),
  Variance_Explained = var_explained_20[1:25]
)

ggplot(df, aes(x = PC, y = Variance_Explained)) +
  geom_bar(stat = "identity", fill = "#EE799F") +
  geom_text(aes(label = sprintf("%.3f", Variance_Explained)), vjust = -0.5, size = 2) +
  theme_minimal(base_size = 14) +
  labs(
    
    x = "Principal Component",
    y = "Proportion of Variance Explained"
  ) +
  theme(
    panel.grid.major.y = element_line(color = "grey90"),
    axis.text.x = element_text(angle = 0, hjust = 1)
  ) + theme_blank() +
  ylim(0, max(df$Variance_Explained) * 1.10)

# ggsave("./figures/paper_figures/FigureS3.png", width = 2200, height = 1200, units = "px", type = "cairo")
```
:::

# Trait vs PCs

:::panel-tabset
## Legends for plotting

**function to plot legends for fig 3**

```{r}
legend_region <- ggplot(fgmerge, aes(x = PC2, y = BudBreak_cGDD_plasticity,
                                           shape=Region) ) +
              geom_point( size=3, alpha=0.9) +
              theme_bw(base_size = 11, base_family = "Times") +
              ylab(expression(atop("Plasticity of bud break",paste("(cGDD)")))) +
              theme(axis.text=element_text(size=14), 
                    axis.title=element_text(size=12),
                    legend.text=element_text(size=rel(.8)), 
                    strip.text = element_text(size=30), 
                    legend.position = "bottom")

legend_ancestry <- ggplot(fgmerge, aes(x = PC2, y = BudBreak_cGDD_plasticity,color = q2) ) +
              geom_point( size=3, alpha=0.9) +
              
              scale_color_gradient("Ancestry",low = "grey95",high = "#251714FF") +
              geom_smooth(aes(x=PC2, y=BudBreak_cGDD_plasticity),method="lm",
                          se=FALSE,formula = y ~ x,
                          inherit.aes = FALSE, lwd=1, colour="#009E73") +
              theme_bw(base_size = 11, base_family = "Times") +
              ylab(expression(atop("Plasticity of bud break",paste("(cGDD)")))) +
              theme(axis.text=element_text(size=14), 
                    axis.title=element_text(size=12),
                    panel.background = element_rect(fill = "transparent"),
                    legend.text=element_text(size=rel(.8)), 
                    strip.text = element_text(size=30),
                    legend.position = "bottom")




# ggsave("./figures/Qpc/bud_break_plasticity_vs_PC2_ancestry_coloring.png", width = 1200, height = 2000, units = "px", type = "cairo")
```

**function to plot legends for supplementary figure**

```{r}
# function to reposition the legend in one of the empty facets
require(lemon) # required package

# code from: https://stackoverflow.com/questions/54438495/shift-legend-into-empty-facets-of-a-faceted-plot-in-ggplot2

shift_legend3 <- function(p) {
    pnls <- cowplot::plot_to_gtable(p) %>% gtable::gtable_filter("panel") %>%
      with(setNames(grobs, layout$name)) %>% purrr::keep(~identical(.x,zeroGrob()))

    if( length(pnls) == 0 ) stop( "No empty facets in the plot" )

    lemon::reposition_legend( p, "center", panel=names(pnls) )
}
```

## Data Wrangling

**Trait data**

```{r}
# set up the data frame
fgmerge_LA_traits <- fgmerge[,c("Family","Budbreak_2020" ,"BudSet_2019","BudSet_2020","Growth",
                                "BudBreak_cGDD_plasticity","BudSet_2019_plasticity",
                                "BudSet_2020_plasticity","Growth_plasticity","TotalDry_plasticity",
                                "RootShoot_plasticity","SideBranch_MainStem_plasticity",
                                "Delta13Carbon_plasticity","Delta15Nitrogen_plasticity",
                                "PercentN_plasticity","CN_ratio_plasticity","TotalDry","RootShoot",
                                "SideBranch_MainStem","delta13C","delta15N","percentN" ,"CN_ratio",
                                "PC1","PC2",#"PC3",
                                "State","Region", "Latitude",
                                "Longitude","Elevation","q1","q2")]

# traits to plot 
trait_id <- c("BudSet_2019","Budbreak_2020","Growth","TotalDry")

fgmerge_LA_traits_long <- fgmerge_LA_traits %>%
  pivot_longer(cols = all_of(trait_id), names_to = "Trait", values_to = "TraitValue")

require(plyr)
fgmerge_LA_traits_long$Figure_axis_text <- revalue(fgmerge_LA_traits_long$Trait,
                                               c(#SideBranch_MainStem = "Side branch:Main stem",
                                               # RootShoot = "Root:Shoot",
                                               TotalDry = "Total dry biomass",
                                               # TotalDry_plasticity = "Total dry biomass (p)",
                                               Growth = "Height growth",
                                               # Growth_plasticity="Height growth (p)",
                                               #CN_ratio = "C:N",
                                               # PercentN = "Percent N",
                                               # Delta15Nitrogen = "delta 15N",
                                               # delta13C = "delta 13C",
                                               # RootShoot_plasticity = "Root:Shoot",
                                               # SideBranch_MainStem_plasticity = "Side branch:Main stem",
                                               # Delta13Carbon_plasticity = "delta 13C",
                                               # Delta15Nitrogen_plasticity = "delta 15N",
                                               # PercentN_plasticity = "Percent N",
                                               # CN_ratio_plasticity = "C:N",
                                               # BudBreak_2020_plasticity = "Bud break (2020)",
                                               # BudSet_2020_plasticity = "Bud set (2020)",
                                               # BudSet_2019_plasticity = "Bud set (2019)",
                                               Budbreak_2020 = "Bud break (2020)",
                                               # BudSet_2020 = "Bud set (2020)",
                                               BudSet_2019 = "Bud set (2019)"))

fgmerge_LA_traits_long$Figure_axis_text <- factor(fgmerge_LA_traits_long$Figure_axis_text,
                                                  levels = c("Bud set (2019)","Bud break (2020)",
                                                    "Height growth","Total dry biomass"))
```

**Plasticity trait**

```{r}
# plasticity on PC1
# plasticity to plot 
pc1_plasticity_id <- c("BudBreak_cGDD_plasticity")

fgmerge_plasticity_pc1_long <- fgmerge_LA_traits %>%
  pivot_longer(cols = all_of(pc1_plasticity_id), names_to = "Trait", values_to = "TraitValue")

fgmerge_plasticity_pc1_long$Figure_axis_text <- revalue(fgmerge_plasticity_pc1_long$Trait,
                                               c(
                                               BudBreak_cGDD_plasticity = "Bud break (2020)"))

# plasticity on PC2
# plasticity to plot 
pc2_plasticity_id <- c("Growth_plasticity","TotalDry_plasticity")

fgmerge_plasticity_pc2_long <- fgmerge_LA_traits %>%
  pivot_longer(cols = all_of(pc2_plasticity_id), names_to = "Trait", values_to = "TraitValue")

fgmerge_plasticity_pc2_long$Figure_axis_text <- revalue(fgmerge_plasticity_pc2_long$Trait,
                                               c(
                                               TotalDry_plasticity = "Total dry biomass",
                                               Growth_plasticity="Height growth"))

fgmerge_plasticity_pc2_long$Figure_axis_text <- factor(fgmerge_plasticity_pc2_long$Figure_axis_text,
                                                  levels = c("Height growth","Total dry biomass"))
```

**Trait vs PC plots**

```{r}
# PC 1 plot

PC1_traits <- fgmerge_LA_traits_long[fgmerge_LA_traits_long$Trait==c("BudSet_2019","Budbreak_2020"),]

pc1_trait_plot <- ggplot(data = PC1_traits, aes(y = TraitValue, x = PC1)) +
                  geom_point(aes(shape = Region, color = q2), size = 4, alpha = 0.9) +
                  # Range wide
                  # geom_smooth(method = "lm", se = F, color = "#56B4E9", fill = "#009E73", alpha = 0.3) +
                  geom_smooth(aes(x=PC1, y=TraitValue),method="lm",
                                          se=FALSE,formula = y ~ x,
                                          inherit.aes = FALSE, lwd=1, colour="#009E73") +
                  # Admixed (Core + Margin)
                  # geom_smooth(data = growth_clim_long %>% filter(Region != "Edge"), 
                  #             method = "lm", se = TRUE, color = "maroon", fill = "darkred", alpha = 0.2) +
                  scale_color_gradient("Ancestry", low = "grey95",high = "#251714FF") +
                  facet_wrap(~ Figure_axis_text, scales = "free_y", 
                             nrow = 2,strip.position = "top", dir = "v") +
                    theme_bw(base_size = 11, base_family = "Times") +
                  theme(axis.text = element_text(size = 14),
                        axis.title = element_text(size = 12),
                        panel.background = element_blank(),
                        legend.background = element_blank(),
                        panel.grid = element_blank(),
                        # plot.background = element_blank(),
                        legend.text = element_text(size = rel(0.8)),
                        strip.text = element_text(size = 15),
                        legend.position = "none",
                        # strip.background = element_rect(fill=NA), #
                        legend.box = "horizontal") +
                  labs(y = "", x = "PC1") 
            

# PC 2 plot
PC2_traits <- fgmerge_LA_traits_long[fgmerge_LA_traits_long$Trait==c("Growth","TotalDry"),]

pc2_trait_plot <- ggplot(data = PC2_traits, aes(y = TraitValue, x = PC2)) +
                  geom_point(aes(shape = Region, color = q2), size = 4, alpha = 0.9) +
                  # Range wide
                  # geom_smooth(method = "lm", se = F, color = "#56B4E9", fill = "#009E73", alpha = 0.3) +
                  geom_smooth(aes(x=PC2, y=TraitValue),method="lm",
                                          se=FALSE,formula = y ~ x,
                                          inherit.aes = FALSE, lwd=1, colour="#009E73") +
                  # Admixed (Core + Margin)
                  # geom_smooth(data = growth_clim_long %>% filter(Region != "Edge"), 
                  #             method = "lm", se = TRUE, color = "maroon", fill = "darkred", alpha = 0.2) +
                  scale_color_gradient("Ancestry", low = "grey95",high = "#251714FF") +
                  facet_wrap(~ Figure_axis_text, scales = "free_y", 
                             nrow = 2,strip.position = "top", dir = "v") +
                    theme_bw(base_size = 11, base_family = "Times") +
                  theme(axis.text = element_text(size = 14),
                        axis.title = element_text(size = 12),
                        panel.background = element_blank(),
                        legend.background = element_blank(),
                        panel.grid = element_blank(),
                        # plot.background = element_blank(),
                        legend.text = element_text(size = rel(0.8)),
                        strip.text = element_text(size = 15),
                        legend.position = "none",
                        # strip.background = element_rect(fill=NA), #
                        legend.box = "horizontal") +
                  labs(y = "", x = "PC2") 

# save figure
# pdf - 12 x 5 in
# png - 1500 x 656 px

# redo the plot in cowplot for ggsave to work

# ggsave("./figures/Figure4.png", width = 3000, height = 1000,dpi=300,
#        limitsize = T,
#        units = "px", type = "cairo")
# 
# ggsave("./figures/Figure4.pdf", width = 30, height = 25,dpi=300,
#        units = "cm")

```

**Plasticity vs PC plots**

```{r}
# PC 1 plot
pc1_plasticity_plot <- ggplot(data = fgmerge_plasticity_pc1_long, aes(y = TraitValue, x = PC1)) +
                        geom_point(aes(shape = Region, color = q2), size = 4, alpha = 0.9) +
                        # Range wide
                        # geom_smooth(method = "lm", se = F, color = "#56B4E9", fill = "#009E73", alpha = 0.3) +
                        geom_smooth(aes(x=PC1, y=TraitValue),method="lm",
                                                se=FALSE,formula = y ~ x,
                                                inherit.aes = FALSE, lwd=1, colour="#009E73") +
                        # Admixed (Core + Margin)
                        # geom_smooth(data = growth_clim_long %>% filter(Region != "Edge"), 
                        #             method = "lm", se = TRUE, color = "maroon", fill = "darkred", alpha = 0.2) +
                        scale_color_gradient("Ancestry", low = "grey95",high = "#251714FF") +
                        facet_wrap(~ Figure_axis_text, scales = "free_y", 
                                   nrow = 2,strip.position = "top", dir = "v") +
                          theme_bw(base_size = 11, base_family = "Times") +
                        theme(axis.text = element_text(size = 14),
                              axis.title = element_text(size = 12),
                              panel.background = element_blank(),
                              legend.background = element_blank(),
                              panel.grid = element_blank(),
                              # plot.background = element_blank(),
                              legend.text = element_text(size = rel(0.8)),
                              strip.text = element_text(size = 15),
                              legend.position = "none",
                              # strip.background = element_rect(fill=NA), #
                              legend.box = "horizontal") +
                        labs(y = "", x = "PC1") 
            

# PC 2 plot
pc2_plasticity_plot <- ggplot(data = fgmerge_plasticity_pc2_long, aes(y = TraitValue, x = PC2)) +
                        geom_point(aes(shape = Region, color = q2), size = 4, alpha = 0.9) +
                        # Range wide
                        # geom_smooth(method = "lm", se = F, color = "#56B4E9", fill = "#009E73", alpha = 0.3) +
                        geom_smooth(aes(x=PC2, y=TraitValue),method="lm",
                                                se=FALSE,formula = y ~ x,
                                                inherit.aes = FALSE, lwd=1, colour="#009E73") +
                       
                        scale_color_gradient("Ancestry", low = "grey95",high = "#251714FF") +
                        facet_wrap(~ Figure_axis_text, scales = "free_y", 
                                   nrow = 2,strip.position = "top", dir = "v") +
                          theme_bw(base_size = 11, base_family = "Times") +
                        theme(axis.text = element_text(size = 14),
                              axis.title = element_text(size = 12),
                              panel.background = element_blank(),
                              legend.background = element_blank(),
                              panel.grid = element_blank(),
                              # plot.background = element_blank(),
                              legend.text = element_text(size = rel(0.8)),
                              strip.text = element_text(size = 15),
                              legend.position = "none",
                              # strip.background = element_rect(fill=NA), #
                              legend.box = "horizontal") +
                        labs(y = "", x = "PC2") 
```

## Paper Figure 3

```{r message=FALSE, warning=FALSE}
#| fig-height: 10
#| fig-width: 15
panel_3A <- qpc_plot
panel_3B <- PC1_PC2_plot
panel_3C <- plot_grid(pc1_trait_plot,pc2_trait_plot)


bottom_row <- cowplot::plot_grid(panel_3B, panel_3C, labels = c('B', 'C'), label_size = 12,
                                 rel_widths = c(1,1.2))
top_row <- cowplot::plot_grid(panel_3A + guides(color = guide_legend(nrow = 1, byrow = TRUE, 
                              keywidth = unit(3, "lines"), 
                              keyheight = unit(1, "lines"),
                              label.vjust = 0.5)))

plot_fig <- cowplot::plot_grid(top_row, bottom_row, labels = c('A', ''), label_size = 12, ncol = 1,
                   rel_heights = c(1,1.4))




legend1 = cowplot::get_plot_component(qpc_plot + scale_x_discrete(guide = guide_axis(check.overlap = TRUE)), 'guide-box-top', return_all = TRUE)
legend2 = cowplot::get_plot_component(legend_ancestry, 'guide-box-bottom', return_all = TRUE)
legend3 = cowplot::get_plot_component(legend_region, 'guide-box-bottom', return_all = TRUE)

cowplot::plot_grid(plot_fig, 
                   cowplot::plot_grid(legend2,legend3, nrow = 1),
                   nrow = 2, rel_heights = c(0.6,0.1),
                   rel_widths = c(1,1))




# ggsave("./figures/paper_figures/Figure3.png", width = 3800, height = 3012,dpi=300,
#        limitsize = F,
#        units = "px", type = "cairo")
# 
# ggsave("./figures/paper_figures/Figure3.pdf", width = 30, height = 25,dpi=300,
#        units = "cm")
```

## Paper Figure 4 Supplementary  
- Plasticity plots  
```{r message=FALSE, warning=FALSE}
#| fig-height: 7
#| fig-width: 10
plot3S_legend <- PC1_PC2_plot + (theme(legend.position = "right",legend.box = "horizontal"))

panel_3S_leftTOP <- pc1_plasticity_plot
panel_3S_leftBOTTOM <- get_plot_component(plot3S_legend, 'guide-box-right', return_all = TRUE)
  
panel_3S_leftPANEL <- cowplot::plot_grid(panel_3S_leftTOP, 
                                         plot_grid(panel_3S_leftBOTTOM, nrow = 1), 
                                         labels = c('', ''), 
                                         label_size = 12, 
                                         ncol = 1,
                                         rel_heights = c(1.12,1))

panel_3S_rightPANEL <- pc2_plasticity_plot

plot_platicity_fig <- cowplot::plot_grid(panel_3S_leftPANEL, panel_3S_rightPANEL, 
                               labels = c('', ''), label_size = 12, ncol = 2)


plot_platicity_fig




# ggsave("./figures/paper_figures/Figure4S.png", width = 2500, height = 2000,dpi=300,
#        limitsize = F,
#        units = "px", type = "cairo")
# 
# ggsave("figures/paper_figures/FigureS4.pdf", width = 20, height = 15,dpi=300,
#        units = "cm")
```


:::
